### 2019-11-05 analysis/model.Rmd
gemma

---
title: "Model"
author: "Siming Zhao"
date: "10/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model
 $$Y = \tilde{X}\gamma + G\theta + \epsilon$$
$Y$: quantitative traits, $N$ x 1 vector, $N$ is number of individuals  
$\tilde{X}$: imputed gene expression (standardized). $N$ x $J$ matrix, $J$ is number of genes.  
$G$: genotype (standardized). $N$ x $M$ matrix, $M$ is number of SNPs.  
$\gamma$: effect sizes for $\tilde{X}$,  $\gamma_j$, effect size for gene $j$.

$$ \gamma_j \sim N(0, \sigma_\gamma^2)$$
$\theta$: effect sizes for $G$,  $\theta_m$, effect size for SNP $m$.  
$$ \theta_m \sim N(0, \sigma_\theta^2)$$

## Model inference
The above model is equivalent to this: 
$$Y \sim N(0,\sigma_\gamma^2XX^T + \sigma_\theta^2GG^T + \sigma_e^2I) $$
This is REML with multiple (more than 2) variance components. Use HE regression function from GEMMA software.