### 2019-11-05 analysis/simulation-WTCCC.Rmd

---
title: "simulation-WTCCC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### WTCCC data

### Fusion results

## Simulate quantitative phenotype
Simulate different $PVE_{SNP}$, $PVE_{gene}$.

$$PVE_{SNP} = \frac{M\sigma_\theta^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
$$PVE_{gene} = \frac{J\sigma_\gamma^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
Scenario 1: $PVE_{SNP}=0.2$, $PVE_{gene}=0.2$