### 2019-11-07 analysis/simulation-WTCCC.Rmd
simulate description

---
title: "simulation-WTCCC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Use WTCCC data obtained from Peter on CRI.

### Simulate cis-expression 
We simulate cis- expression based on the following model:

$$ \tilde{X} = G\alpha + \xi$$
For each gene, we sample $K$ eQTLs for this gene, for eQTL $k$, we have
$$ \alpha_k \sim N(0,\sigma_\alpha^2) $$
$$ \xi \sim N(0,1) $$
Then we have the heritability of the gene:

$$\begin{aligned}
h^2_{eQTL} &= \frac{var(G\alpha)}{var(G\alpha)+var(\xi)} \\
&= \frac{\Sigma_kvar(G_k) \alpha_k^2}{\Sigma_kvar(G_k) \alpha_k^2 + var(\xi)}\\
&=  \frac{\Sigma_k\alpha_k^2}{\Sigma_k\alpha_k^2 + var(\xi)}\\
&\approx \frac{\Sigma_k\sigma_\alpha^2}{\Sigma_k\sigma_\alpha^2 + var(\xi)} \\
&=\frac{K\sigma_\alpha^2}{1+K\sigma_\alpha^2}
\end{aligned}$$

Here, we use scaled genotype data, so $var(G)=1$. We also have $\alpha^2_k \approx E(\alpha_k^2)=var(\alpha_k^2)=\sigma_\alpha^2$. Usually, $K$ ranges from 1 to 5, let $\sigma_\alpha = 0.3$, then $h^2_{eQTL}$ ranges from 0.08 to 0.31, this is consistent with gene cis-heritability in reality.

### Simulate quantitative phenotype
Simulate different $PVE_{SNP}$, $PVE_{gene}$.

$$PVE_{SNP} = \frac{M\sigma_\theta^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
$$PVE_{gene} = \frac{J\sigma_\gamma^2}{J\sigma_\gamma^2 + M\sigma_\theta^2 + \sigma_e^2}$$
Scenario 1: $PVE_{SNP}=0.2$, $PVE_{gene}=0.2$