### 2020-03-13 analysis/simulation-WTCCC2.Rmd
v1.0 simulate expression using separate data 20191127

---
title: "Simulation-WTCCC-2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Use WTCCC data obtained from Peter on CRI.

The difference between this analysis and [previous run](simulation-WTCCC.html): used 500 samples from `cad.RData` from WTCCC to simulate expression, then train expression model using `cv.glmnet`. We impute expression in using `bd.RData` from WTCCC.

Simulation of phenotype: use the same weights as used in simulate expression using the `'cad.RData` dataset.

The genotype data and other settings are the same.

## Simulate cis-expression 
We simulate cis-expression based on the following model:

$$ \tilde{X} = G\alpha + \xi$$
For each gene, we sample $L$ eQTLs for this gene, for eQTL $k$, we have
$$ \alpha_k \sim N(0,\sigma_\alpha^2) $$
$$ \xi \sim N(0,1) $$
Then we have the heritability of the gene:

$$\begin{aligned}
h^2_{eQTL} &= \frac{var(G\alpha)}{var(G\alpha)+var(\xi)} \\
&= \frac{\Sigma_kvar(G_k) \alpha_k^2}{\Sigma_kvar(G_k) \alpha_k^2 + var(\xi)}\\
&=  \frac{\Sigma_k\alpha_k^2}{\Sigma_k\alpha_k^2 + var(\xi)}\\
&\approx \frac{\Sigma_k\sigma_\alpha^2}{\Sigma_k\sigma_\alpha^2 + var(\xi)} \\
&=\frac{K\sigma_\alpha^2}{1+K\sigma_\alpha^2}
\end{aligned}$$

Here, we use scaled genotype data, so $var(G)=1$. We also have $\alpha^2_k \approx E(\alpha_k^2)=var(\alpha_k^2)=\sigma_\alpha^2$. Usually, $K$ ranges from 1 to 5, let $\sigma_\alpha = 0.3$, then $h^2_{eQTL}$ ranges from 0.08 to 0.31, this is consistent with gene cis-heritability in reality.

## Simulate quantitative phenotype
We simulate phenotype followin

$$\begin{align}
PVE_{SNP} = &\frac{var(G\theta)}{var(G\theta) + var(\tilde{X}\gamma) + \sigma_e^2}\\ 
\approx &\frac{M\sigma_\theta^2}{M\sigma_\theta^2 + \Sigma_jvar(\tilde{X_j})\gamma_j^2+ \sigma_e^2}\\ 
\approx &\frac{M\sigma_\theta^2}{M\sigma_\theta^2 + Jvar(\tilde{X})\sigma_\gamma^2+ \sigma_e^2}\\   
\end{align}$$

$$
PVE_{expr} \approx \frac{Jvar(\tilde{X})\sigma_\gamma^2}{M\sigma_\theta^2 + Jvar(\tilde{X})\sigma_\gamma^2+ \sigma_e^2}$$

Here, $var(\tilde{X})$ is the cis-heratbility of gene expression. $M$ and $J$ are numbers of causal SNP and gene respectively.

Thus in order to get desired $PVE_{SNP}$ and $PVE_expr$, we set $\sigma_\theta^2$ and $\sigma_\gamma^2$ based on the following formula:

$$\sigma_\theta^2 = \frac{PVE_{SNP}}{M(1-PVE_{SNP} - PVE_{expr})}$$
$$\sigma_\gamma^2 = \frac{PVE_{expr}}{Jvar(\tilde{X})(1-PVE_{SNP} - PVE_{expr})}$$

```{r}
simudir <- "/home/simingz/causalTWAS/simulations/simulation_WTCCC_20191127/"
show_res <- function(simutag){
  load(Sys.glob(paste0(simudir,simutag,"*phenotype.Rd")))
  outdf1 <- data.frame(truth = c(sigma_gamma^2*J.c, sigma_theta^2*M.c , 1))
  row.names(outdf1) <- c("sigma_gamma^2","sigma_theta^2","sigma_e^2")
  res <- readLines(paste0(simudir, simutag, "_VC/output/", simutag, ".log.txt"))
  outdf1$est <- as.numeric(strsplit(res[24], "  ")[[1]][2:4])
  outdf1$est.se <- as.numeric(strsplit(res[25], "  ")[[1]][2:4])
  print(outdf1)
  outdf2 <- data.frame("est"= c(as.numeric(strsplit(res[20], "  ")[[1]][2:3]), as.numeric(strsplit(res[22], "=")[[1]][2])))
  row.names(outdf2) <- c("PVE.expr","PVE.snp","PVE.total")
  outdf2$est.se <- c(as.numeric(strsplit(res[21], "  ")[[1]][2:3]), as.numeric(strsplit(res[23], "=")[[1]][2]))
  print(outdf2)
}
```

## Simulation 1: $M$ = No. of all SNPs, $J$ = No. all genes
We simulate quantitative phenotype under several scenarios. First we make all genes with imputed gene expression as causal genes and all SNPs as causal SNPs. This matches our polygenic version of the model.

### 1.1 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.1$
```{r 1.1}
show_res("S1.1")
```

### 1.2 $PVE_{SNP}\approx 0.4$, $PVE_{expr}\approx 0.1$
```{r 1.2}
show_res("S1.2")
```

## Simulation 2: $M$ = No. of all SNPs, $J$ = ~10% all genes (J=2000)

### 2.1 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.1$
```{r 2.1}
show_res("S2.1")
```

### 2.2 $PVE_{SNP} \approx 0.4$, $PVE_{expr} \approx 0.1$
```{r 2.2}
show_res("S2.2")
```
## Simulation 3: $M$ = ~ 1% all SNPs (M=5000), $J$ = No. all genes

### 3.1 $PVE_{SNP} \approx 0.1$, $PVE_{expr} \approx 0.1$ 
```{r 3.1}
show_res("S3.1")
```

### 3.2 $PVE_{SNP} \approx 0.4$, $PVE_{expr} \approx 0.1$
```{r 3.2}
show_res("S3.2")
```