### 2020-03-31 analysis/VEB-boost-fitting.Rmd
sparse model description

---
title: "VEB-boost"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem description

$$Y_{N \times 1}= X_{N \times J}\beta_{J\times1} + G_{N\times M}\theta_{M\times1} + \epsilon_{N\times1}$$
$$ \epsilon \sim MVN(0,I_N\tau^{-1})$$
Spike and slab prior for $\beta_j$:
$$\beta_j|\pi,\tau \sim (1- \pi_\beta)\delta_0 + \pi_\beta N(0, \sigma_\beta^2/\tau)$$
Spike and slab prior for $\theta_m$:
$$\theta_m | \tau \sim (1- \pi_\theta)\delta_0 + \pi_\theta N(0, \sigma_\theta^2/\tau)$$
We want to estimate $\omega=\{\tau,\pi_\beta, \pi_\theta, \sigma_\beta^2,\sigma_\theta^2\}$