### 2020-03-31 analysis/sparse_model.Rmd

---
title: "Sparse model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sparse model for expression and genotype
 $$Y = \tilde{X}\beta + G\theta + \epsilon$$
$Y$: quantitative traits, $N$ x 1 vector, $N$ is number of individuals, the columns are centered.
$\tilde{X}$: the genetic component of gene expression, $N$ x $J$ matrix, $J$ is number of genes, the columns are centered.
$G$: genotype (standardized). $N$ x $M$ matrix, $M$ is number of SNPs.

The error term $\epsilon$ has a normal distribution,

$$\epsilon|\tau \sim N(0, \tau^{-1})$$

We use the spike and slab prior for gene expression effect size $\beta_j$ (effect size for gene $j$).
$$\beta_j|\pi,\tau \sim (1- \pi_\beta)\delta_0 + \pi_\beta N(0, \sigma_\beta^2/\tau)$$
Here $\delta_0$ is point mass at 0 ($\delta_0(\beta_j)=1$ if $\beta_j=0$, otherwise $\delta_0(\beta_j)=0$ ).

We use the spike and slab prior for SNP effect size $\theta_m$ (effect size for SNP $m$).
$$ \theta_m | \tau \sim (1- \pi_\theta)\delta_0 + \pi_\theta N(0, \sigma_\theta^2/\tau)$$

We use the following prior distributions for the hyperparameters $\omega = (\pi_\beta,\pi_\theta, \tau, \sigma_\beta^2,\sigma_\theta^2)$: 
$$\tau \sim Gamma(\kappa_1, \kappa_2)$$

$$log(\pi_\theta) \sim U(log(1/M), log(1))$$
$$log(\pi_\beta) \sim U(log(1/J), log(1))$$
We define $h_{SNP}$ and $h_{expr}$ as follows: 
$$
h_{SNP} := \frac{M\sigma_\theta^2}{M\sigma_\theta^2 + Jvar(\tilde{X})\sigma_\beta^2+ 1}$$
$$
h_{expr} := \frac{Jvar(\tilde{X})\sigma_\beta^2}{M\sigma_\theta^2 + Jvar(\tilde{X})\sigma_\beta^2+ 1}$$

Instead of assigning priors for $\sigma_\beta^2$ and $\sigma_\theta^2$, we assign uniform priors for $h_{SNP}$ and $h_{expr}$:

$$ h_{SNP} \sim U(0,1), h_{expr} \sim U(0,1) $$

## Inference
We are interested in the hyperparameters $\omega = (\pi_\beta, \pi_\theta, \tau, \sigma_\beta^2,\sigma_\theta^2)$. We introduce two vectors of binary indicators ($\gamma_\beta$ = $\{\gamma_{\beta1}, \gamma_{\beta2}, ..., \gamma_{\beta J}\} \in \{0,1\}^J$ and $\gamma_\theta$ = $\{\gamma_{\theta1}, \gamma_{\theta2}, ..., \gamma_{\theta M}\} \in \{0,1\}^M$) that indicates whether the corresponding $\beta_j$ or $\theta_m$ is non-zero.

$$\gamma_{\beta j} \sim Bernoulli(\pi_\beta)$$
$$\gamma_{\theta m} \sim Bernoulli(\pi_\theta)$$
$$ \beta_j|\gamma_{\beta j}=1 \sim  N(0, \sigma_\beta^2/\tau), \beta_j|\gamma_{\beta j}=0 \sim  \delta_0$$
$$ \theta_m|\gamma_{\theta m}=1 \sim  N(0, \sigma_\theta^2/\tau), \theta_m|\gamma_{\theta m}=0 \sim  \delta_0$$
The posterior distrition is

$$P(\pi_\beta, \pi_\theta, \tau, h_{expr}, h_{SNP}, \gamma_\beta, \gamma_\theta|y) \propto P(y|\pi_\beta, \pi_\theta, \tau, h_{expr}, h_{SNP}, \gamma_\beta, \gamma_\theta) P(h_{expr})P(h_{SNP})P(\gamma_\beta|\pi_\beta)P( \gamma_\theta|\pi_\theta)P(\pi_\theta)P(\pi_\beta)$$