### 2020-04-16 analysis/mr.ash.Rmd
veb-boost-repo

---
title: "mr.ash"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mr.ash.alpha); # package implementing proposed methods in the paper
library(Matrix); library(mvtnorm) # miscellaneous packages for matrices
library(glmnet); library(ncvreg); library(L0Learn); # penalized regression methods
library(varbvs); library(susieR); library(BGLR); # Bayesian methods
library(ggplot2); library(cowplot); # plotting
```

## Install mr.ash
Following https://github.com/stephenslab/mr.ash.alpha

## mr.ash examples
```{r workflowr}
setwd("/home/simingz/causalTWAS/mr.ash/mr-ash-workflow")
source('code/method_wrapper.R')
source('code/sim_wrapper.R')
data = simulate_data(n = 100, p = 200, s = 5,
                     design = "indepgauss", # "equicorrgauss",
                     signal = "normal", # "lap", "t", etc.
                     pve = 0.5)
str(data)

method_list  = c("mr.ash","varbvs","bayesb","blasso","susie","enet","lasso","ridge","scad2","mcp2","l0learn")
fit.function = get(paste0("fit.", method_list[1], sep = ""))

fit  = fit.function(data$X, data$y, data$X.test, data$y.test)
```

## mr.ash papers
```{r workflowr-paper}
setwd("/home/simingz/causalTWAS/mr.ash/mr-ash-workflow")
filelist = paste("analysis/", list.files("analysis", pattern = "*_Sourcecode.Rmd"), sep = "")
for (i in 1:length(filelist)) {
  workflowr::wflow_publish(filelist[i])
  cat("file",i,": ",'"',filelist[i],'"'," is published\n",sep = "")
}

filelist = paste("analysis/", list.files("analysis", pattern = "*_Plot.Rmd"), sep = "")
for (i in 1:length(filelist)) {
  workflowr::wflow_publish(filelist[i])
  cat("file",i,": ",'"',filelist[i],'"'," is published\n",sep = "")
}
```