### 2020-05-20 analysis/VEB-boost-fitting.Rmd
test mr.ash

---
title: "VEB-boost"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic idea of veb_boost

You want to add up 2 linear regressions, which we might call “learners”.
A learner is, roughly, a fucntion, that learns a relationshipo between Y and some predictors X (here the Xs are different for your 2 different learners) Let’s write this $Y=L1(X1) + L2(X2) + E$.

The idea is that you can fit this model iteratively, by first applying $L1$, and then applying $L2$, each time applying it to the appropriate residuals.So you apply $L1$ to learn a relationship between $R=Y-L2$ (residuals) and $X1$, and then $L2$ to learn a relationship between $R=Y-L1$ and $X2$ and you just iterate….

## Our model

Model:

$$Y = \mu_1 + \mu_2 + \epsilon$$
$$ \mu_1= X_{N \times J}\beta_{J\times1}$$
$$ \mu_2 = G_{N\times M}\theta_{M\times1}$$
$$ \epsilon \sim MVN(0,\sigma^2I_N) $$
Priors:

ASH prior for scaled $\beta_j$ and $\theta_m$
$$\beta_j | g_1, \sigma \sim g^1_\sigma(\cdot)$$
ASH prior for $\theta_m$:
$$\theta_m | g_2, \sigma \sim g^2_\sigma(\cdot)$$

## Input required by veb_boost

take $\mu1$ as the example, $\mu2$ works similarly.

* first moment of posterior variational approximation ($E_q(\mu_i)\in R^N$)

$$E_q(\mu_1) = X\bar{\beta}$$
$\bar{\beta}$ is posterior mean of $\beta$ given by `mr.ash` function.

* second moment of posterior variational approximation ($E_q(\mu^2_i)\in R^N_+$)

$$E_q(\mu_{1}^2) = \sum_{j=1}^J{x^2_j E_q(\beta^2)} + [E_q(\mu_1)]^2 =\sum_{j=1}^J\sum_{k=1}^K{x^2_j\phi_k \frac{\sigma^2 d_j^2 \sigma_k^2}{1+ d_j^2 \sigma_k^2}} + [E_q(\mu_1)]^2$$
where $$d_j=x^T_jx_j= \left\lVert x_j\right\rVert^2$$.

$d_j, \phi, \sigma, \sigma_k$ are estimates given by `mr.ash` function.

* KL-divergence between variational posterior (q) and prior (p).

Shall we just use ELBO from mr.ash?