### 2020-05-20 analysis/simulation_simpleveb-boost.Rmd

---
title: "Simple simulations for a simplified version of web-boost"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mr.ash.alpha)
source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")
source("~/causalTWAS/causal-TWAS/code/simple_vebboost.R")
```

## Simulation of data

20 blocks:

* Each block has either gene or SNP effect
* Each block has 99 SNPs and 1 gene. Each gene is linear sum of the previous 3 SNPs. 
* Each block, either the gene or the last eQTL has non-zero effect on trait

The first 4 blocks have gene effect.

```{r generate data}
set.seed(1)
N <- 4000
nblocks <- 20
block.size <- 100
p <- nblocks * block.size
n.eQTL <- 3  # number of eQTLs per gene
sigma.eQTL <- 0.5 # eQTL effect size
sigma.SNP <- 0.1 # effect size of causal SNP on trait
sigma.gene <- 0.1 # effect size of causal gene on trait
X <- matrix(rep(0,0), nrow=N, ncol=0)
gamma.gene <- rep(0, nblocks) # indicator of genes
gamma.gene[1:4] <- 1 
beta <- numeric(0)
SNP.idx <- numeric(0)
for (i in 1:nblocks) {
  # sample SNP data
  X.block.SNP <- matrix(rnorm(N*(block.size-1)), nrow=N, ncol=block.size-1)
  SNP.idx <- c(SNP.idx, 1:(block.size-1) + (i-1)*block.size)

  # generate gene data: use the previous few SNPs as eQTL
  effects.eQTL <- rnorm(n.eQTL, 0, sigma.eQTL)
  X.block.gene <- X.block.SNP[, (block.size - n.eQTL):(block.size - 1)] %*% effects.eQTL
  X.block = cbind(X.block.SNP, X.block.gene)
  X <- cbind(X, X.block)

  # sample beta
  if (gamma.gene[i] == 1) { # gene effect in this block
    beta.SNP <- rep(0, block.size - 1)
    beta.gene <- rnorm(1, 0, sigma.gene)
  } else { # SNP effect in this block
    beta.SNP <- c(rep(0, block.size - 2), rnorm(1, 0, sigma.SNP))
    beta.gene <- 0
  }
  beta.block <- c(beta.SNP, beta.gene)
  beta <- c(beta, beta.block)
}
sigma.e <- 1
y <- X %*% beta + rnorm(N, 0, sigma.e)
gene.idx <- (1:nblocks) * block.size
```

## Run mr.ash

```{r plot functions}
summary_mr.ash <- function(fit){
  cat("pi1 = ", 1-fit$pi[[1]], "\n")
  pve <- get_pve(fit)
  cat("pve : ", pve, "\n")
}

plot_beta <- function(beta,beta.pm, ...){
  plot( beta, pch=19, col ="darkgreen", ...)
  points(beta.pm, pch =19, col = "red")
  legend("topright", legend=c("true beta", "posterior mean"),
       col=c("darkgreen", "red"), pch=19)
}

```

```{r run mr.ash}
fit <- mr.ash(X, y, method="caisa")
summary_mr.ash(fit)

plot_beta(beta[gene.idx], fit$beta[gene.idx], main = "beta for gene effect")
plot_beta(beta[SNP.idx], fit$beta[SNP.idx], main = "beta for SNP effect")
```

## Run a simplified version of veb-boost

### start with gene

```{r run mr.ash2, gene}
X.gene <- X[,gene.idx]
X.SNP <- X[, SNP.idx]
fit <- mr.ash2(X.gene, X.SNP, y)

print("for gene effect: ")
summary_mr.ash(fit$fit1)
plot_beta(beta[gene.idx], fit$fit1$beta, main = "beta for gene effect")

print("for SNP effect: ")
summary_mr.ash(fit$fit2)
plot_beta(beta[SNP.idx], fit$fit2$beta, main = "beta for SNP effect")
```

### start with SNP

```{r run mr.ash2, snp}
X.gene <- X[,gene.idx]
X.SNP <- X[, SNP.idx]
fit <- mr.ash2(X.SNP, X.gene, y)

print("for gene effect: ")
summary_mr.ash(fit$fit2)
plot_beta(beta[gene.idx], fit$fit2$beta, main = "beta for gene effect")

print("for SNP effect: ")
summary_mr.ash(fit$fit1)
plot_beta(beta[SNP.idx], fit$fit1$beta, main = "beta for SNP effect")
```