### 2020-07-03 analysis/simulation-mr.ash2s-ukbchr22-gtex.adipose.Rmd
mr.ash2s lasso beta init

---
title: "Test a simpler version of veb-boost performance mr.ash2s on ukbiobank data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We test the performance of mr.ash2s on ukbiobank data (chr22, n sample = 20,000). 
We use weights from Fusion website, using [Adipose (subcutaneous) p < 0.01](http://gusevlab.org/projects/fusion/weights/GTEx.Adipose_Subcutaneous.P01.tar.bz2).
Total features > 8,000. On chr22: 240.

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})

source("~/causalTWAS/causal-TWAS/code/fit_mr.ash.R")

summary_mr.ash <- function(fit){
  cat("pi1 = ", 1-fit$pi[[1]], "\n")
  pve <- get_pve(fit)
  cat("pve = ", pve, "\n")
}

summary_mr.ash2 <- function(g.fit, s.fit, phenores, pipcut = 0.1){
  e.b <- rep(0, length(g.fit$beta))
  e.b[phenores$param$idx.cgene] <- phenores$param$e.beta
  cat("gene expression effect: \n")
  summary_mr.ash(g.fit)

  s.b <- rep(0, length(s.fit$beta))
  s.b[phenores$param$idx.cSNP] <- phenores$param$s.theta
  cat("snp effect: \n")
  summary_mr.ash(s.fit)
}

plot_pip <- function(dt, ld = F){

  if (isTRUE(ld)) {
    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', 
                   mode = 'markers', color = ~ max_r2, 
                   text = ~ paste("Name: ", name, 
                                  "\n", ifcausal,
                                  "\nmax R^2: ", round(max_r2, digits = 2), 
                                  "\nMax pair: ", max_pair))
  } else {
    pal <- c("salmon", "darkgreen")
    pal <- setNames(pal, c("Causal", "Non causal"))

    fig <- plot_ly(dt, x = ~p0, y = ~ PIP, type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))
  }

  fig <- fig %>% layout(yaxis = list(title = "PIP"))
  return(fig)
}

plot_beta <- function(dt){
  fig <- plot_ly(dt, x = ~p0, y = ~ TrueBeta, type = 'scatter', mode = 'markers', name ="Truth", text = ~paste("Name: ", name))
  fig <- fig %>% add_trace(y = ~ EstBeta, name = 'Estimated',mode = 'markers')

  fig <- fig %>% layout(yaxis = list(title = "beta"))
  return(fig)
}

plot_p <- function(dt){
  pal <- c("salmon", "darkgreen")
  pal <- setNames(pal, c("Causal", "Non causal"))

  fig <- plot_ly(dt, x = ~p0, y = ~ PVALUE , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

  fig <- fig %>% layout(yaxis = list(title = "p value"))
  return(fig)
}

plot_susie_pip <- function(dt) {
   pal <- c("salmon", "darkgreen")
   pal <- setNames(pal, c("Causal", "Non causal"))

   fig <- plot_ly(dt, x = ~ p0, y = ~ susiePIP , type = 'scatter', mode = 'markers', color = ~ ifcausal, colors = pal, text = ~paste("Name: ", name))

   fig <- fig %>% layout(yaxis = list(title = "SUSIE PIP"))
}

plot_all <- function(mrashres.gene, gwasres.gene, mrashres.snp, gwasres.snp, susieres, ldres){
  a <- read.table(mrashres.gene , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  ld <- read.table(ldres, header =T)
  a <- merge(a, ld, by = "name", all.x = T)

  su <- read.table(susieres, header =T)
  colnames(su)[colnames(su) == "pip"] <-  "susiePIP"
  a <- merge(a, su, by = "name", all.x = T)

  p <- fread(gwasres.gene, header =T)
  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
  dt.gene <- merge(a, p, by = "name", all.x = T)

  a <- read.table(mrashres.snp , header =T)
  a[a$ifcausal == 1, "ifcausal"] = "Causal"
  a[a$ifcausal == 0, "ifcausal"] = "Non causal"

  p <- fread(gwasres.snp, header =T)
  p[,"PVALUE"] <- -log10(p[,"PVALUE"])
  colnames(p)[colnames(p) == "MARKER_ID"] <-  "name"
  dt.snp <- merge(a, p, by = "name", all.x = T)

  fig1 <- plot_pip(dt.gene, ld = T)  
  fig2 <- plot_beta(dt.gene) 
  fig3 <- plot_p(dt.gene) 
  fig4 <- plot_susie_pip(dt.gene) 
  fig5 <- plot_pip(dt.snp) 
  fig6 <- plot_beta(dt.snp) 
  fig7 <- plot_p(dt.snp) %>% layout( plot_bgcolor='#EFF6FC')

  fig <- subplot(fig1, fig2, fig3, fig4, fig5, fig6, fig7, nrows=7, shareX = T, titleY = T)

  fig <- fig %>% layout(autosize = F, width = 600, height = 900, margin = 0.02,
    xaxis = list(title = "Genomic position"))

  fig <- fig %>% toWebGL()
  fig
}

```

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200616/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200616/"
```

```{r getfile function}
get_files <- function(tag, tag2){
  par <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt")
  gmrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".expr.txt")
  ggwas <- paste0(outputdir, tag, ".exprgwas.txt.gz")
  smrash <- paste0(outputdir, tag, "-mr.ash2s.", tag2, ".snp.txt")
  sgwas <- paste0(outputdir, tag, ".snpgwas.txt.gz")
  susie <- paste0(susiedir, tag, ".", tag2, ".L3-susieres.txt")
  rsq <- paste0(outputdir, tag, ".LD.txt")
  return(tibble::lst(par, gmrash, ggwas, smrash, sgwas, susie, rsq))
}
```

## Simulation 1
MR.ASH2s results (start from gene):
```{r sim 1 gene, out.width = "98%", out.height="95%"}
tag <- "20200616-1"
tag2 <- "expr-res"

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)
```

MR.ASH results (start from SNP):
```{r sim 1 snp, eval = F}
tag <- "20200616-1"
tag2 <- "snp-res"

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)
```

## Simulation 2

MR.ASH results (start from gene):
```{r sim 2 gene, out.width = "98%", out.height="95%"}
tag <- "20200616-2"
tag2 <- "expr-res"

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)
```
MR.ASH results (start from snp):
```{r sim 2 snp, out.width = "98%", out.height="95%"}
tag <- "20200616-2"
tag2 <- "snp-res"

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)
```

## Simulation 3

MR.ASH results (start from gene):
```{r sim 3 gene, out.width = "98%", out.height="95%"}
tag <- "20200616-3"
tag2 <- "expr-res"

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)
```
MR.ASH results (start from snp):
```{r sim 3 snp, out.width = "98%", out.height="95%"}
tag <- "20200616-3"
tag2 <- "snp-res"

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)
```

## Simulation 4

MR.ASH results (start from gene):
```{r sim 4 gene, out.width = "98%", out.height="95%"}
tag <- "20200616-4"
tag2 <- "expr-res"

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)
```

MR.ASH results (start from snp):
```{r sim 4 snp, out.width = "98%", out.height="95%"}
tag <- "20200616-4"
tag2 <- "snp-res"

param <- read.table(paste0(outputdir, tag, "-mr.ash2s.", tag2, ".param.txt"), header = T, row.names = 1)
print(param)
```

## Simulation 5

MR.ASH results (start from gene):
```{r sim 5 gene, out.width = "98%", out.height="95%"}
tag <- "20200616-5"
tag2 <- "expr-res"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

MR.ASH results (start from SNP): very similar to start from gene.
```{r sim 5 snp, out.width = "98%", out.height="95%", eval = F}
tag <- "20200616-5"
tag2 <- "snp-res"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

## Simulation 6

MR.ASH results (start from gene):
```{r sim 6 gene, out.width = "98%", out.height="95%"}
tag <- "20200616-6"
tag2 <- "expr-res"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

MR.ASH results (start from SNP): very similar to start from gene
```{r sim 6 snp, out.width = "98%", out.height="95%", eval = F}
tag <- "20200616-6"
tag2 <- "snp-res"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

## Simulation 7
Use lasso trained weight model from FUSION twas. This makes eQTLs (SNP weights != 0) more sparse (<20/gene) compared to BSLMM/BLUP (a few hundred/per gene).

Use beta.init = lasso for mr.ash.

MR.ASH results (use lasso for all):
```{r sim 7 lasso, out.width = "98%", out.height="95%", eval = T}
tag <- "20200616-7"
tag2 <- "lasso"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

MR.ASH results (use lasso for just for SNPs):
```{r sim 7 lassosnp, out.width = "98%", out.height="95%", eval = T}
tag <- "20200616-7"
tag2 <- "lassoSNP"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

## Simulation 8
Use lasso trained weight model from FUSION twas. This makes eQTLs (SNP weights != 0) more sparse (<20/gene) compared to BSLMM/BLUP (a few hundred/per gene).

Use beta.init = lasso for mr.ash.

MR.ASH results (use lasso for all):
```{r sim 8 lasso, out.width = "98%", out.height="95%", eval = T}
tag <- "20200616-8"
tag2 <- "lasso"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```

MR.ASH results (use lasso for just for SNPs):
```{r sim 8 lassosnp, out.width = "98%", out.height="95%", eval = T}
tag <- "20200616-8"
tag2 <- "lassoSNP"

f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$susie, f$rsq)
```