### 2020-08-05 analysis/simulation-multi-ukbchr17to22-gtex.adipose.Rmd
move code-  plot functions

---
title: "Summarize twas: chr17 to 22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F)
```

Run simulation 8 times for ukb chr 17 to chr 22 combined. SNPs are downsampled to 1/10, eQTLs defined by FUSION-TWAS (Adipose/GTEx) lasso effect size > 0 were kept, p= 86k, n = 20k.

```{r plot functions}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
library(kableExtra)
source("analysis/summarize_twas_plots.R")
```
```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/sa2_default/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/sa2_default/"
tags <- paste0('20200721-1-', c(2, 4:9))
tagglob <- '20200721-1-*'
tagextr <- '20200721-1-\\d+'
tag2s <- c('zeroes-es', 'zerose-es', 'lassoes-es','lassoes-se')
```

# Mr.ash2 parameter estimation

Results for 10 simulations runs, using different initiate and update strategy

## NULL; expr-snp; expr-snp

```{r param 1}
show_param(tags, tag2s[1])
```

## NULL; snp-expr; expr-snp

```{r param 2}
show_param(tags, tag2s[2])
```

## lasso; expr-snp; expr-snp

```{r param 3}
show_param(tags, tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r param 4}
show_param(tags, tag2s[4])
```

# Regional mr.ash2s PIP overview
Take simulation 1 (NULL; expr-snp; expr-snp) as examples. We use region size *500kb* and PIP cut off at *0.2* for SUSIE.

```{r rpip 1, fig.height=4, fig.width=10}
chrom = 18
f <- get_files(tag= "20200721-1-2" , tag2 = tag2s[1])
allchr <- read.table(f[["rpip"]], header = T)
a <- allchr[allchr["chrom"]==chrom,]
print(paste("plot for chr", chrom))
par(mar=c(5, 4, 4, 6) + 0.1)
with(a, plot(p0, rPIP, col ='salmon', xlab = "position", ylab= "Sum of PIP", type = 'h', lwd = 2))
par(new = T)
with(a, plot(p0, nCausal, pch =19, col = "darkgreen",axes = FALSE, bty = "n", xlab = "", ylab = ""))
axis(side = 4)
mtext(side = 4, line = 3, 'No. causal signals')
legend("topleft",
       legend=c("Mr.ASH PIP", "# Causal"),
       lty=c(1,0), pch=c(NA, 19), col=c("salmon", "darkgreen"))
grid()
```

# PIP calibration
We run 50 simulations and combine results.

## NULL; expr-snp; expr-snp 
```{r cali 1, fig.height=4, fig.width=14}
tag2 = "zeroes-es"
tags_ext <- Reduce(intersect, get_tags(tagglob, tagextr, tag2 = tag2)['gsusie'])
res <- caliPIP_plot(tags = tags_ext, tag2 = tag2)
```
```{r cali 1fdr, fig.height=4, fig.width=4}
tags_ext2 <- Reduce(intersect, get_tags(tagglob, tagextr, tag2 = tag2)['gsusie'])
caliFDR_plot(tags = tags_ext2, tag2 = tag2)
```

## Lasso; expr-snp; expr-snp 
```{r cali 2, fig.height=4, fig.width=14}
caliPIP_plot(tags = tags_ext, tag2 = "lassoes-es")
```

```{r cali 2fdr, fig.height=4, fig.width=4}
caliFDR_plot(tags = tags_ext2, tag2 = "lassoes-es")
```

# PIP scatter plot
mr.ash2s PIP vs. susie PIP.

## NULL; expr-snp; expr-snp 
```{r sca 1, fig.height=3, fig.width=10}
scatter_plot_PIP(tags, tag2s[1])
```

## NULL; snp-expr; expr-snp

```{r sca 2, fig.height=3, fig.width=10}
scatter_plot_PIP(tags, tag2s[2])
```

## lasso; expr-snp; expr-snp

```{r sca 3, fig.height=3, fig.width=10}
scatter_plot_PIP(tags, tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r sca 4, fig.height=3, fig.width=10}
scatter_plot_PIP(tags, tag2s[4])
```

# ROC curve

## NULL; expr-snp; expr-snp 
```{r roc 1, fig.height=4, fig.width=4}
tags <- paste0('20200721-1-', c(2,4:9))
ROC_plot(tags, tag2s[2])
```

## NULL; snp-expr; expr-snp

```{r roc 2, fig.height=4, fig.width=4}
ROC_plot(tags, tag2s[2])
```

## lasso; expr-snp; expr-snp

```{r roc 3, fig.height=4, fig.width=4}
ROC_plot(tags, tag2s[3])
```

## lasso; expr-snp; snp-expr

```{r roc 4, fig.height=4, fig.width=4}
ROC_plot(tags, tag2s[4])
```