### 2020-09-03 analysis/simulation-susieP-ukbchr17to22-gtex.adipose.Rmd

---
title: "Test mr.ash2s on ukbiobank data, chr17-22"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message= F, warning = F)
```

We test the performance of mr.ash2s on ukbiobank data (chr17 to 22, n sample = 20,000). 
We use weights from Fusion website, using [Adipose (subcutaneous) p < 0.01](http://gusevlab.org/projects/fusion/weights/GTEx.Adipose_Subcutaneous.P01.tar.bz2).
Total features > 8,000. On chr17 to 22: ~1500.

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20200721/"
```

```{r source file}
source("analysis/visual_twas_plots.R")
```

## Simulation 1
## NULL; expr-snp; expr-snp
```{r sim 1-4, out.width = "98%", out.height="95%"}
tag <- "20200721-1-4"
tag2 <- "zeroes-es"
f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
```
```{r 1-4-17, out.width = "98%", out.height="95%"}
chrom <- "17"
cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
```

```{r 1-4-19 gene, out.width = "98%", out.height="95%"}
chrom <- "19"
cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
```

```{r sim 1-8, out.width = "98%", out.height="95%"}
tag <- "20200721-1-8"
f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
```
```{r 1-8-20, out.width = "98%", out.height="95%"}
chrom <- "20"
cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
```

```{r sim 1-9, out.width = "98%", out.height="95%"}
tag <- "20200721-1-9"
f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
```
```{r 1-9-19, out.width = "98%", out.height="95%"}
chrom <- "19"
cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
```