### 2020-09-14 analysis/simulation-susieI-ukbchr17to22-gtex.adipose.Rmd

---
title: "Test susieI using chr17 to 22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
```
```{r libs}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
library(kableExtra)
source("analysis/summarize_twas_plots.R")
```

```{r des, child = here::here("analysis/description.Rmd")}
```

```{r dirs}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20200813/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20200813/"
mrashdir <- "~/causalTWAS/simulations/simulation_ashtest_20200721/"
tag <- '20200813-1-1'
Niter <- 30
```

# Regional minimum p value distribution

* minimum TWAS p value of genes for each region:

```{r rpip twas, fig.height=6, fig.width=6}
simtag <- "20200721-1-1"
rpipf <- paste0(simdatadir, simtag, ".pgenerprank.txt")

a <- read.table(rpipf, header = T)
a.c <- a[a$nCausal > 0, ]
ax <- pretty(0:max(-log10(a$rpmin)), n = 30)

par(mfrow=c(2,1))
h1 <- hist(-log10(a$rpmin), breaks = 100, xlab = "PIP", main = "min P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
```

* minimum TWAS p value of genes and SNPs for each region:

```{r rpip twas2, fig.height=6, fig.width=6}
rpipf <- paste0(simdatadir, simtag, ".p2rprank.txt")

a <- read.table(rpipf, header = T)
a.c <- a[a$nCausal > 0, ]
ax <- pretty(0:max(-log10(a$rpmin)), n = 30)

par(mfrow=c(2,1))
h1 <- hist(-log10(a$rpmin), breaks = 100, xlab = "PIP", main = "min P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
```

# MR.ASH regional pip sum distribution

* Use mr.ash for gene only:

```{r rpip 1, fig.height=6, fig.width=6, eval=F}
rpipf <- paste0(simdatadir, simtag, "-mr.ash.mrashrPIP.txt")

a <- read.table(rpipf, header = T)
a.c <- a[a$nCausal > 0, ]
ax <- pretty(0:max(a$rPIP), n = 30)

par(mfrow=c(2,1))
h1 <- hist(a$rPIP, breaks = 100, xlab = "PIP", main = "PIP distribution-all", col = "grey"); grid()
h2 <- hist(a.c$rPIP, breaks = h1$breaks, xlab = "PIP", main = "PIP distribution-causal", col = "salmon");grid()
```

* MR.ash2s for both genes and SNPs using the same data

```{r rpip 2, fig.height=6, fig.width=6}
rpipf <- paste0(simdatadir, simtag, "-mr.ash2s.lassoes-es.mrash2srPIP.txt")

a <- read.table(rpipf, header = T)
a.c <- a[a$nCausal > 0 ,]
ax <- pretty(0:max(a$rPIP), n = 30) # Make a neat vector for the breakpoints

par(mfrow=c(2,1))
h1 <- hist(a$rPIP, breaks = 100, xlab = "PIP", main = "PIP distribution-all", col = "grey"); grid()
h2 <- hist(a.c$rPIP, breaks = h1$breaks, xlab = "PIP", main = "PIP distribution-causal", col = "salmon");grid()
```

# Parameter estimation results

Results: Each row shows parameter estimation results from 5 simulation runs with similar settings (i.e. pi1 and PVE for genes and SNPs). each row has two plots, one for gene pi1 estimation, one for enrichment (gene pi1/snp pi1). Results from each run were represented by one dot, dots with the same color come from the same run. horizontal dash lines: simulation truth, `mr.ash2`, mr.ash for gene and SNPs; `susietruth`, the truth in selected regions that were used to run susie iteractively (susieI).

## Regions: causal

* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3, and also require at least one causal gene or SNP is present, also, at least one gene.

* Susie run parameters: `L=1`, `null_weight = 0`, didn't update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform.

```{r param functions}
show_param <- function(mrashfs, susieIfs, susieIfs2){
  param <- do.call(rbind, lapply(mr.ashfs, function(x) t(read.table(x))[2:1,]))

  pars <- cbind(param[seq(1,nrow(param), 2), 2], param[seq(1,nrow(param), 2), 4], param[seq(1,nrow(param), 2), 1], param[seq(2,nrow(param), 2), 1], param[seq(1,nrow(param), 2), 3], param[seq(2,nrow(param), 2), 3])
  colnames(pars) <- c("PVE.gene_truth", "PVE.SNP_truth", paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("truth", "mr.ash2"), sep = ""))

  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec, 1), tail(prior.SNP_rec,1))}))

  param.s.truth <- do.call(rbind, lapply(susieIfs2, function(x) {
    a <- read.table(x, header = T);
    c(nrow(a[a$ifcausal == 1 & a$type == "gene", , drop = F])/ nrow(a[a$type == "gene", , drop = F]),
      nrow(a[a$ifcausal == 1 & a$type == "SNP", , drop = F])/ nrow(a[a$type == "SNP", , drop = F]))
    }))

  pars.s <- cbind(param.s.truth, param.s)[, c(1,3,2,4)]
  colnames(pars.s) <-  paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("susietruth", "susieI"), sep = "")

  df <- cbind(tags, format(pars, digits = 4), format(pars.s, digits =4))
  rownames(df) <- NULL
  return(df)

  # df %>% 
  # kable("html", escape = F) %>%
  # kable_styling("striped", full_width = F) %>%
  #  row_spec(c(1:5, 11:15), background = "#FEF3B9") %>%
  # scroll_box(width = "100%", height = "600px", fixed_thead = T)
}

plot_param <- function(df, ...){
  df <- apply(df[ , 2:ncol(df)], 2, function(x) as.numeric(x))
  m <- cbind(df[,"pi1.gene_mr.ash2"], 1:nrow(df), 1 + 1:nrow(df)/nrow(df)/3)
  st <-  cbind(df[,"pi1.gene_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]
  dfp <- rbind(m,st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "gene pi1", xaxt = "n", xlab="", xlim = c(0.8, 3.5), frame.plot=FALSE, ylim = c(0, max(dfp[,1],t) *1.05), ...)
  axis(side=1, at=1:3, labels = FALSE, tick = F)
  text(x=1:3, 0, labels = c("mr.ash2", "susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h=t, lty = 2, col= "salmon", lwd=1.5)
  grid()

  m <- cbind(df[,"pi1.gene_mr.ash2"]/df[,"pi1.SNP_mr.ash2"], 1:nrow(df), 1 + 1:nrow(df)/nrow(df)/3)
  st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"]/df[,"pi1.SNP_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]/df[1,"pi1.SNP_truth"]
  dfp <- rbind(m,st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, max(dfp[,1],t, 150) *1.05))
  axis(side=1, at=1:3, labels = FALSE, tick = F)
  text(x=1:3, 0, labels = c("mr.ash2", "susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h= t, lty = 2, col= "darkgreen", lwd=1.5)
  grid()
}
```

```{r param, fig.height=8, fig.width=6}
tags <- paste(rep(1:3, each = 5), 1:5, sep = "-")
mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
susieIfs <- paste0(outputdir, "20200813-", tags, ".susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20200813-", tags, ".1.txt")

df <- show_param(mrashfs, susieIfs, susieIfs2)
par(mfrow = c(3,2))
plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05") 
plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
```

## Regions: mrash gene+SNP (1)

* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3.

* Susie run parameters: `L=2`, initialize `null_weight = 0`, update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform. (We also tried initialize with prior for genes and SNPs as 0.05 and 2.5e-3, results are the same.)

```{r param-6, fig.height=11, fig.width=6}
tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
susieIfs <- paste0(outputdir, "20200813-", tags, ".config6.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config6.susieI.txt")

df <- show_param(mrashfs, susieIfs, susieIfs2)
par(mfrow = c(4,2))
plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
```

## Regions: mrash gene+SNP (2)

* Regions: We use regional PIP of both SNP and gene from mr.ash2s > 0.3.

* Susie run parameters: `L=2`, `null_weight = 0`. Didn't update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform.

```{r param-7, fig.height=11, fig.width=6}
tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
susieIfs <- paste0(outputdir, "20200813-", tags, ".config7.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config7.susieI.txt")

df <- show_param(mrashfs, susieIfs, susieIfs2)
par(mfrow = c(4,2))
plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
```

## Regions: mrash gene only

* Regions: We use regional sum of PIP for genes from mr.ash for genes only > 0.1.

* Susie run parameters: `L=2`. Update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform, `null_weight = 0`.

```{r param-9, fig.height=11, fig.width=6}
tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
susieIfs <- paste0(outputdir, "20200813-", tags, ".config9.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config9.susieI.txt")

df <- show_param(mrashfs, susieIfs, susieIfs2)
par(mfrow = c(4,2))
plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
```

## Regions: TWAS gene only

* Regions: We use regional min p value for genes from twas for genes only, requiring this p value in top 5%.

* Susie run parameters: `L=2`. Update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform, `null_weight = 0`.

```{r param-10, fig.height=11, fig.width=6}
tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
susieIfs <- paste0(outputdir, "20200813-", tags, ".config10.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config10.susieI.txt")

df <- show_param(mrashfs, susieIfs, susieIfs2)
par(mfrow = c(4,2))
plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
```

## Regions: TWAS gene+SNP

* Regions: We use regional min p value for genes or SNPs from twas for genes and snps, requiring this p value in top 5% for genes or 1% for SNPs.

* Susie run parameters: `L=2`. Update `null_weight` in iterations. We initialize with prior for genes and SNPs as uniform, `null_weight = 0`.

```{r param-11, fig.height=11, fig.width=6}
tags <- paste(rep(1:4, each = 5), 1:5, sep = "-")
susieIfs <- paste0(outputdir, "20200813-", tags, ".config11.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config11.susieI.txt")

df <- show_param(mrashfs, susieIfs, susieIfs2)
par(mfrow = c(4,2))
plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
```

## Regions: all (1)

* Regions: All regions.

* Susie run parameters: `L=2`. initialize with `null_weight = 0` and update `null_weight` based on last iteration results. We initialize with prior for genes and SNPs as uniform.

```{r param3, fig.height=11, fig.width=6}
mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
susieIfs <- paste0(outputdir, "20200813-", tags, ".config2.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config2.susieI.txt")

df <- show_param(mrashfs, susieIfs, susieIfs2)
par(mfrow = c(4,2))
plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
```

## Regions: all (2)

* Regions: All regions.

* Susie run parameters: `L=1`. Did not update `null_weight` in each iteration. We initialize with prior for genes and SNPs as uniform.

```{r param4, fig.height=11, fig.width=6}
mr.ashfs <- paste0(mrashdir, "20200721-", tags, "-mr.ash2s.lassoes-es.param.txt")
susieIfs <- paste0(outputdir, "20200813-", tags, ".config3.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20200813-", tags, ".config3.susieI.txt")

df <- show_param(mrashfs, susieIfs, susieIfs2)
par(mfrow = c(4,2))
plot_param(df[1:5,], main = "1-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[6:10,], main = "2-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[11:15,], main = "3-x: gene.pve ~ 0.01, snp.pve ~ 0.05")
plot_param(df[16:20,], main = "4-x: gene.pve ~ 0.05, snp.pve ~ 0.05")
```

## Regions: all (others)
* Susie run parameters: `L=1`. initialize with null_weight = 0 and update null_weight based on last iteration results. We initialize with prior for genes and SNPs as uniform. The prior for genes is underestimated for low power settings (1-x, 2-x), many converged to close to 0, for high power settings it is better. 
* Susie run parameters: `L=5`. initialize with null_weight = 0 and update null_weight based on last iteration results. The prior for gene estimation is fine for all runs, however prior for SNPs overestimated about 10 times.
* We have tried different initiation strategies, but the results remain the same.