### 2020-10-02 analysis/simulation-susieI-ukbchr17to22B-gtex.adipose.Rmd

---
title: "Test susieI using chr17 to 22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
```
```{r libs}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
library(kableExtra)
source("analysis/summarize_twas_plots.R")
```

```{r des, child = here::here("analysis/description.Rmd")}
```

```{r dirs}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20200915/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20200915/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20200915/"
tag <- '20200915-1-1'
Niter <- 30
```

# Regional minimum p value distribution

* minimum TWAS p value of genes for each region:

```{r rpip twas, fig.height=6, fig.width=6}
simtag <- "20200915-1-1"
rpipf <- paste0(outputdir, simtag, ".pgenerprank.txt")

a <- read.table(rpipf, header = T)
a.c <- a[a$nCausal > 0, ]
ax <- pretty(0:max(-log10(a$rpmin)), n = 30)

par(mfrow=c(2,1))
h1 <- hist(-log10(a$rpmin), breaks = 100, xlab = "PIP", main = "min P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
```

* minimum TWAS p value of genes and SNPs for each region:

```{r rpip twas2, fig.height=6, fig.width=6}
rpipf <- paste0(outputdir, simtag, ".p2rprank.txt")

a <- read.table(rpipf, header = T)
a.c <- a[a$nCausal > 0, ]
ax <- pretty(0:max(-log10(a$rpmin)), n = 30)

par(mfrow=c(2,1))
h1 <- hist(-log10(a$rpmin), breaks = 100, xlab = "PIP", main = "min P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(a.c$rpmin), breaks = h1$breaks, xlab = "PIP", main = "min p distribution-causal", col = "salmon");grid()
```

# Parameter estimation results

Results: Each row shows parameter estimation results from 5 simulation runs with similar settings (i.e. pi1 and PVE for genes and SNPs). each row has two plots, one for gene pi1 estimation, one for enrichment (gene pi1/snp pi1). Results from each run were represented by one dot, dots with the same color come from the same run. horizontal dash lines: simulation truth, `mr.ash2`, mr.ash for gene and SNPs; `susietruth`, the truth in selected regions that were used to run susie iteractively (susieI).

```{r param functions}
show_param <- function(phenofs, susieIfs, susieIfs2){
  pars <- do.call(rbind, lapply(phenofs, function(x) {load(read.table(x, header = F, stringsAsFactors = F)[1,1]); 
    c(phenores$param$pve.gene.truth,
      phenores$param$pve.snp.truth,
      length(phenores$param$idx.cgene)/phenores$param$J,
      length(phenores$param$idx.cSNP)/phenores$param$M)}))

  colnames(pars) <- c("PVE.gene_truth", "PVE.SNP_truth", "pi1.gene_truth", "pi1.SNP_truth")

  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec[prior.gene_rec!=0], 1), tail(prior.SNP_rec[prior.SNP_rec!=0],1))}))

  param.s.truth <- do.call(rbind, lapply(susieIfs2, function(x) {
    a <- read.table(x, header = T);
    c(nrow(a[a$ifcausal == 1 & a$type == "gene", , drop = F])/ nrow(a[a$type == "gene", , drop = F]),
      nrow(a[a$ifcausal == 1 & a$type == "SNP", , drop = F])/ nrow(a[a$type == "SNP", , drop = F]))
    }))

  pars.s <- cbind(param.s.truth, param.s)[, c(1,3,2,4)]
  colnames(pars.s) <-  paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("susietruth", "susieI"), sep = "")

  df <- cbind(tags, format(pars, digits = 4), format(pars.s, digits =4))
  rownames(df) <- NULL
  return(df)

  # df %>% 
  # kable("html", escape = F) %>%
  # kable_styling("striped", full_width = F) %>%
  #  row_spec(c(1:5, 11:15), background = "#FEF3B9") %>%
  # scroll_box(width = "100%", height = "600px", fixed_thead = T)
}

plot_param <- function(df, ...){
  df <- apply(df[ , 2:ncol(df)], 2, function(x) as.numeric(x))
  st <-  cbind(df[,"pi1.gene_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "gene pi1", xaxt = "n", xlab="", xlim = c(0.8, 3.5), frame.plot=FALSE, ylim = c(0, max(dfp[,1],t) *1.05), ...)
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=1:2, 0, labels = c( "susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h=t, lty = 2, col= "salmon", lwd=1.5)
  grid()

  st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"]/df[,"pi1.SNP_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]/df[1,"pi1.SNP_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, min(max(dfp[,1],t) *1.05, 150)))
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=1:2, 0, labels = c("susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h= t, lty = 2, col= "darkgreen", lwd=1.5)
  grid()
}

gpip_dist <- function(susiefs, ...){
  dflist <- list()
  for (f in susiefs){
    dflist[[f]] <- read.table(f, header =T , stringsAsFactors = F)
  }
  df <- do.call(rbind, dflist)
  hist(df[df$type == "gene", "susie_pip"], xlab = "gene susie PIP", 
       breaks = 50, ylim = c(0,20), xlim=c(0,1), col = "salmon", ...)
}
```

## Regions: all

* Regions: All regions.

* Susie run parameters: `L=2`. initialize with `null_weight = 0` and update `null_weight` based on last iteration results. We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.01, snp.pve ~ 0.05.

```{r param1, fig.height=6, fig.width=6}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20200915-", tags, "-pheno.txt")
susieIfs <- paste0(outputdir, "20200915-", tags, ".config1.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20200915-", tags, ".config1.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,2))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

```{r pipdist-1,fig.height=3, fig.width=12}
par(mfrow = c(1,2))
gpip_dist(susieIfs2[1:5], main = "low power")
gpip_dist(susieIfs2[6:10], main = "high power")
```