### 2020-10-14 analysis/simulation-susieI-ukbWG-gtex.adipose.Rmd

---
title: "Test susieI using whole genome, 4k ukb samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
```
```{r libs}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
library(kableExtra)
source("analysis/summarize_twas_plots.R")
```

```{r des, child = here::here("analysis/description.Rmd")}
```

```{r dirs}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
tag <- '20201001-1-1'
Niter <- 20
```

# p value distribution

* TWAS p value of genes:

```{r gene twas, fig.height=6, fig.width=8}
simtag <- "20201001-1-1"
exprgwasf <- paste0(simdatadir, simtag, ".exprgwas.txt.gz")
load(paste0(simdatadir, "simu_", simtag, "-pheno.Rd"))
caulist <- list()
for (chrom in 1:22) {
  load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40000.FBM.Rd"))
  load(paste0(simdatadir, "simu_s40000_GTEXadipose-B", chrom, "-cis-expr.Rd"))
  caulist[[chrom]]<- c(exprres$gnames[phenores$batch[[chrom]]$param$idx.cgene], dat$snp[phenores$batch[[chrom]]$param$idx.cSNP,])
}
cau <- unlist(caulist)

a <- read.table(exprgwasf, header = T)
a$ifcausal <- ifelse(a$MARKER_ID %in% cau, 1, 0)
ax <- pretty(0:max(-log10(a$PVALUE)), n = 30)

par(mfrow=c(3,1))
h1 <- hist(-log10(a$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(a[a$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon");grid()

plot(a[a$X.CHROM ==22, ]$BEGIN, -log10(a[a$X.CHROM ==22, ]$PVALUE), col = a[a$X.CHROM ==22, ]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
points(a[a$X.CHROM ==22 & a$ifcausal ==1, ]$BEGIN, -log10(a[a$X.CHROM ==22 & a$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
```

```{r eval=F}
r <- do.call(cbind, lapply(1:22, function(x) phenores$batch[[x]]$Y.g))
GWAA(exprres$expr, matrix(rowSums(r[, c(1:12,14:21)]) + rnorm(N), ncol=1), snpname = exprres$gnames, anno = anno, "temp", family = gaussian, ncore = 1, nSplits = 1, compress = F)
m <- read.table("temp", header =T, comment.char = "")
m$ifcausal <- ifelse(m$MARKER_ID %in% cau, 1, 0)
plot(m[m$X.CHROM ==22, ]$BEGIN, -log10(m[m$X.CHROM ==22, ]$PVALUE), col = m$ifcausal + 1)

snpname = dat$snp[,1]
anno <- cbind(dat$chr, dat$pos, dat$pos) # EPACT format

GWAA(dat$G, r[,13, drop=F], snpname = snpname, anno = anno, "tempsnp", family = gaussian, ncore = 5, nSplits = 100, compress = F)
```

* TWAS p value of SNPs:

```{r snp twas, fig.height=6, fig.width=8}
simtag <- "20201001-1-1"
snpgwasf <- paste0(simdatadir, simtag, ".snpgwas.txt.gz")

b <- read.table(snpgwasf, header = T)
b$ifcausal <- ifelse(b$MARKER_ID %in% cau, 1, 0)
ax <- pretty(0:max(-log10(b$PVALUE)), n = 30)

par(mfrow=c(3,1))
h1 <- hist(-log10(b$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey"); grid()
h2 <- hist(-log10(b[b$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon");grid()

plot(b[b$X.CHROM ==22, ]$BEGIN, -log10(b[b$X.CHROM ==22, ]$PVALUE), col = b$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
points(b[b$X.CHROM ==22 & b$ifcausal ==1, ]$BEGIN, -log10(b[b$X.CHROM ==22 & b$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
```

# Parameter estimation results

Results: Each row shows parameter estimation results from 5 simulation runs with similar settings (i.e. pi1 and PVE for genes and SNPs). each row has two plots, one for gene pi1 estimation, one for enrichment (gene pi1/snp pi1). Results from each run were represented by one dot, dots with the same color come from the same run. horizontal dash lines: simulation truth,  `susietruth`, the truth in selected regions that were used to run susie iteractively (susieI).

```{r param functions}
show_param <- function(phenofs, susieIfs, susieIfs2){
  pars <- do.call(rbind, lapply(phenofs, function(x) {load(x); 
    c(phenores$param$pve.gene.truth,
      phenores$param$pve.snp.truth,
      length(phenores$batch[[1]]$param$idx.cgene)/phenores$batch[[1]]$param$J,
      length(phenores$batch[[1]]$param$idx.cSNP)/phenores$batch[[1]]$param$M)}))

  colnames(pars) <- c("PVE.gene_truth", "PVE.SNP_truth", "pi1.gene_truth", "pi1.SNP_truth")

  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec[prior.gene_rec!=0], 1), tail(prior.SNP_rec[prior.SNP_rec!=0],1))}))

  param.s.truth <- do.call(rbind, lapply(susieIfs2, function(x) {
    a <- fread(x, header = T);
    c(nrow(a[a$ifcausal == 1 & a$type == "gene" ])/ nrow(a[a$type == "gene"]),
      nrow(a[a$ifcausal == 1 & a$type == "SNP"])/ nrow(a[a$type == "SNP"]))
    }))

  pars.s <- cbind(param.s.truth, param.s)[, c(1,3,2,4)]
  colnames(pars.s) <-  paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("susietruth", "susieI"), sep = "")

  df <- cbind(tags, format(pars, digits = 4), format(pars.s, digits =4))
  rownames(df) <- NULL
  return(df)

  # df %>% 
  # kable("html", escape = F) %>%
  # kable_styling("striped", full_width = F) %>%
  #  row_spec(c(1:5, 11:15), background = "#FEF3B9") %>%
  # scroll_box(width = "100%", height = "600px", fixed_thead = T)
}

plot_param <- function(df, ...){
  df <- apply(df[ , 2:ncol(df)], 2, function(x) as.numeric(x))
  st <-  cbind(df[,"pi1.gene_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "gene pi1", xaxt = "n", xlab="", xlim = c(0.8, 3.5), frame.plot=FALSE, ylim = c(0, max(dfp[,1],t) *1.05), ...)
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=2:3, 0, labels = c( "susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h=t, lty = 2, col= "salmon", lwd=1.5)
  grid()

  st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"]/df[,"pi1.SNP_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]/df[1,"pi1.SNP_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, min(max(dfp[,1],t) *1.05, 150)))
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=2:3, 0, labels = c("susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h= t, lty = 2, col= "darkgreen", lwd=1.5)
  grid()
}

gpip_dist <- function(susiefs, ...){
  dflist <- list()
  for (f in susiefs){
    dflist[[f]] <- read.table(f, header =T , stringsAsFactors = F)
  }
  df <- do.call(rbind, dflist)
  hist(df[df$type == "gene", "susie_pip"], xlab = "gene susie PIP", 
       breaks = 50, ylim = c(0,20), xlim=c(0,1), col = "salmon", ...)
}
```

## Regions: all (1)

* Susie run parameters: `L=2`. initialize with `null_weight = 0` and update `null_weight` based on last iteration results. We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.

```{r param1, fig.height=6, fig.width=6}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,2))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

## Regions: causal

* Susie run parameters: `L=2`. initialize with `null_weight = 0` and update `null_weight` based on last iteration results. We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.

```{r param5, fig.height=6, fig.width=6}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config5.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config5.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,2))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```