### 2020-10-20 analysis/simulation-susieI-ukbWG-gtex.adipose_visual.Rmd

---
title: "Test mr.ash2s on ukbiobank data, chr1-22, 40k"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message= F, warning = F)
```
Provide visualization for analysis [here]

```{r output dir}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
outputdir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
susiedir <- "~/causalTWAS/simulations/simulation_susietest_20201001/"
```

```{r source file}
source("analysis/visual_twas_plots.R")
```

## Simulation- high power

```{r sim 1-4, out.width = "98%", out.height="95%"}
tag <- "20201001-1-4"
f <- get_files(tag= tag, tag2 = tag2)
param <- read.table(f$par, header = T, row.names = 1)
print(param)
```
```{r 1-4-17, out.width = "98%", out.height="95%"}
chrom <- "17"
cat("simulation run: ", tag, "-", tag2, "; chr: ", chrom)
plot_all(f$gmrash, f$ggwas, f$smrash, f$sgwas, f$gsusie, f$ssusie, f$rsq, chrom = chrom)
```