### 2020-11-13 analysis/simulation-susieI-ukbWG-gtex.adipose_s40.22.Rmd
pip cali

---
title: "Test susieI using whole genome, s40.22 samples"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,  message= F, warning = F)
```
```{r libs}
library(mr.ash.alpha)
library(data.table)
suppressMessages({library(plotly)})
library(tidyr)
library(plyr)
library(stringr)
source("analysis/summarize_twas_plots.R")
```

# Analysis description

```{r parameters}
n.ori <- 40000 # number of samples
n <-  22542
p <- 656321 # number of SNPs
J <- 8021 # number of genes
```

The genotype data we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples. We then filtered samples based on relatedness, ethics and other qc metrics, that ended up with n = `r as.character(n)` samples.  We use SNP genotype data from chr 1 to chr 22 combined from UKB. SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back. This ends up with p = `p as.charater(p)` SNPs.

Our analysis consists of the following steps:

1. Build expression predictors using another expression-genotype dataset.

The one we used in this analysis is GTEx Adipose tissue v7 dataset. This dataset contains ~ 380 samples. [FUSION/TWAS](http://gusevlab.org/projects/fusion/) were used to train expression model and we used their lasso results. SNPs included in eQTL anlaysis are restricted to cis-locus 500kb on either side of the gene boundary. eQTLs are defined as SNPs with abs(effectize) > 1e-8 in lasso results.

2. Impute expression.

We impute gene expression for our genotype data using expression models obtained from step 1. There are `r as.character(J)` genes with expression model. We imputed expression from genotypes using the expression predictors.

3. Define and select regions

Next, the analysis is done at the “region” level, which is 500kb bins along the genome. We also used LDetect to define regions. We are exploring several ways to select regions that contain true signals, e.g. based on regional sum of mr.ash PIP for genes/SNPs, region smallest TWAS p value for gene/SNPs, or regional bayes factors, etc.

4. Run susie iteratively
We then run susie for each of these regions. So the features of SuSiE are: SNPs and "genes" (not cis-eQTLs of that gene). We use the same prior for all SNPs and another prior for all "genes" when running SUSIE. In some settings, we also run SUSIE with null weight, which is calculated as `1- prior.SNP * n.SNP - prior.gene * n.gene`. We obtain the PIP for SNPs and gene in the region. After we run susie for all regions (one iteration), we take the average of all SNP PIPs as the prior of SNPs for the next iteration and similarly for the prior for genes.

5. We obtain PIP for genes from the last iteration as results.

```{r dirs}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
simtag <- '20201001-1-3'
source('~/causalTWAS/causal-TWAS/code/gwas.R')
source('~/causalTWAS/causal-TWAS/code/ld.R')
source('~/causalTWAS/causal-TWAS/code/qqplot.R')
```

```{r causal}
exprgwasf <- paste0(simdatadir, simtag, ".exprgwas.txt.gz")
load(paste0(simdatadir, "simu_", simtag, "-pheno.Rd"))
caulist <- list()
for (chrom in 1:22) {
  load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40.22.FBM.Rd"))
  load(paste0(simdatadir, "simu_s40000_GTEXadipose-B", chrom, "-cis-expr.Rd"))
  caulist[[chrom]]<- c(exprres$gnames[phenores$batch[[chrom]]$param$idx.cgene], dat$snp[phenores$batch[[chrom]]$param$idx.cSNP,])
}
cau <- unlist(caulist)
```
# Power estimation

We use gene.pve ~ 0.1, snp.pve ~ 0.5.

* For SNPs, we use $\pi_1 = 2.5e-3$, variance for effect size ~ $0.03^2$, power at 5e-8 p value cutoff:
```{r pow func} 
pow <- function(total, n, beta, cutp){
  rec <- rep(0, total)
  for (i in 1:total){
    x <- rnorm(n)
    y <- x * rnorm(1, sd = beta) + rnorm(n, sd = sqrt(2.5))
    lm.s <- lm(y~x)
    pv <- summary(lm.s)$coefficients[2,4]
    rec[i] <- pv
  }
  length(rec[rec < cutp])/length(rec)
}
```

```{r snp}
load("data/power_s40.22.Rd")
total <- 1e3
n <- 22542
#p1 <- pow(total, n, 0.0276, 5e-8)
print(p1)
```

* For genes, under low power setting, $\pi_1 = 0.05$, variance for effect size ~ $0.025^2$, power at 1e-5 cutoff:

```{r gene1}
#p2 <- pow(total, n, 0.025, 1e-5)
print(p2)
```

For genes, under high power setting, $\pi_1 = 0.02$, variance for effect size ~ $0.045^2$, power at 1e-5 cutoff:

```{r gene2}
#p3 <- pow(total, n, 0.045, 1e-5)
print(p3)
#save(p1,p2,p3, file = "data/power_s40.22.Rd")
```

# p value distribution

* TWAS p value of genes:

```{r gene twas, fig.height=8, fig.width=8}
chrom <- 1
a <- read.table(exprgwasf, header = T)
a$ifcausal <- ifelse(a$MARKER_ID %in% cau, 1, 0)
ax <- pretty(0:max(-log10(a$PVALUE)), n = 30)

par(mfrow=c(3,1))
h1 <- hist(-log10(a$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey", xlim= c(3,20), ylim =c(0,50)); grid()
h2 <- hist(-log10(a[a$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon", xlim= c(3,20), ylim =c(0,50));grid()

cat("number of genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5, ]))
cat("number of causal genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5 & a$ifcausal ==1, ]))

plot(a[a$X.CHROM ==chrom, ]$BEGIN, -log10(a[a$X.CHROM ==chrom, ]$PVALUE), col = a[a$X.CHROM ==chrom, ]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
points(a[a$X.CHROM ==chrom & a$ifcausal ==1, ]$BEGIN, -log10(a[a$X.CHROM ==chrom & a$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
grid()
```

* qq plot for genes
```{r qqgene, fig.height= 5, fig.width=5}
gg_qqplot(a$PVALUE) +
  theme_bw(base_size = 24) +
  theme(
    axis.ticks = element_line(size = 0.5),
    panel.grid = element_blank()
    # panel.grid = element_line(size = 0.5, color = "grey80")
  )
```

*  p value of SNPs (GWAS):

```{r snp twas, fig.height=8, fig.width=8}
snpgwasf <- paste0(simdatadir, simtag, ".snpgwas.txt.gz")

b <- fread(snpgwasf, header = T)
b$ifcausal <- ifelse(b$MARKER_ID %in% cau, 1, 0)
ax <- pretty(0:max(-log10(b$PVALUE)), n = 30)

par(mfrow=c(3,1))
h1 <- hist(-log10(b$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey", xlim= c(3,20), ylim =c(0,100)); grid()
h2 <- hist(-log10(b[b$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon", xlim= c(3,20), ylim =c(0,100));grid()

cat("number of SNPs < 5e-8: ", nrow(b[b$PVALUE < 5e-8,]))

plot(b[b$X.CHROM ==chrom, ]$BEGIN, -log10(b[b$X.CHROM ==chrom, ]$PVALUE), col = b[b$X.CHROM ==chrom,]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
points(b[b$X.CHROM ==chrom & b$ifcausal ==1, ]$BEGIN, -log10(b[b$X.CHROM ==chrom & b$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
grid()
```

* qq plot for SNPs
```{r qqsnp, fig.height= 5, fig.width=5}
gg_qqplot(b$PVALUE) +
  theme_bw(base_size = 24) +
  theme(
    axis.ticks = element_line(size = 0.5),
    panel.grid = element_blank()
    # panel.grid = element_line(size = 0.5, color = "grey80")
  )
```

# TWAS FDP

```{r PIPplot functions}

library(ggplot2)
library(cowplot)
library(plotrix)

.exob <- function(pips, ifcausal, mode = c("PIP", "FDR")){
  a_bin <- cut(pips, breaks= seq(0, 1, by=0.1))
  if (mode == "PIP") {
    ex = c(by(pips, a_bin, FUN = mean))
    ob = c(by(ifcausal, a_bin, FUN = mean))
  } else if (mode == "FDR"){
    ex = c(by(pips, a_bin, FUN = mean))
    ob = 1 - c(by(ifcausal, a_bin, FUN = mean))
  }
  return(list("expected" = ex, "observed" = ob))
}

dot_plot = function(df, main) {
        ggplot(df, aes(x=mean_pip, y=observed_freq)) + 
          geom_errorbar(aes(ymin=observed_freq-se, ymax=observed_freq+se), colour="black", size = 0.5, width=.01) +
          geom_point(size=1.5, shape=21, fill="#002b36") + # 21 is filled circle
          xlab("Expected") +
          ylab("Observed") +
          coord_cartesian(ylim=c(0,1), xlim=c(0,1)) +
          geom_abline(slope=1,intercept=0,colour='red', size=0.2) +
          ggtitle(main) +
          expand_limits(y=0) +                        # Expand y range
          theme_cowplot()}

cp_plot <- function(pips, ifcausal, runtag = NULL, mode = c("PIP", "FDR"), main = mode[1], ...){

  # ifcausal:0,1, runtag: for adding std.
  if (is.null(runtag)){
     se = 0 
  } else{
    dflist <- list()
    for (rt in unique(runtag)){
      pips.rt <- pips[runtag == rt]
      ifcausal.rt <- ifcausal[runtag == rt]
      dflist[[rt]] <- .exob(pips.rt, ifcausal.rt, mode = mode)
    }
    mean_pip <- colMeans(do.call(rbind, lapply(dflist, '[[', "expected")))
    observed_freq <- colMeans(do.call(rbind, lapply(dflist, '[[', "observed")))
    se <- apply(do.call(rbind, lapply(dflist, '[[', "observed")), 2, plotrix::std.error)
  }

  df <- data.frame("mean_pip" = mean_pip, "observed_freq"= observed_freq, "se" = se)
  dot_plot(df, main = main)

  #plot(Expected, Observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main, ...)
  #lines(x = c(0,1), y = c(0,1), col ="grey", lty = 2)
}

caliPIP_plot <- function(pipfs, format = "susie", main = "PIP"){
  df <- NULL
  for (i in 1:length(pipfs)) {
    res <- fread(pipfs[i], header = T)
    res <- data.frame(res[res$type  =="gene", ])
    res$runtag <- i
    if (format == "susie"){
       res <- rename(res, c("susie_pip" = "pip"))
    } else if (format == "SER"){
       res <- rename(res, c("SERpip" = "pip"))
    }
    res <- res[complete.cases(res),]
    df <- rbind(df, res)
  }

  fig <- cp_plot(df$pip, df$ifcausal, df$runtag, mode ="PIP", main = main)
  return(fig)
}

# pipfs is just for ifcausal info
caliFDP_plot <- function(gwasfs, pipfs,  format= "susie",  main = "FDP"){
  df <- NULL
  for (i in 1:length(pipfs)) {
    pipres <- fread(pipfs[i], header = T)
    pipres <- data.frame(pipres[pipres$type  =="gene", ])
    pipres$runtag <- i
    if (format == "susie"){
       pipres <- rename(pipres, c("susie_pip" = "pip"))
    } else if (format == "SER"){
       pipres <- rename(pipres, c("SERpip" = "pip"))
    }
    gwasres <- read.table(gwasfs[i], header = T)
    gwasres <- rename(gwasres, c("MARKER_ID" = "name"))
    gwasres$FDR <- p.adjust(gwasres$PVALUE, method = "fdr")
    res <- merge(gwasres, pipres, by = "name", all = T)
    res <- res[complete.cases(res),]
    df <- rbind(df, res)
  }
  fig <- cp_plot(df$FDR, df$ifcausal, df$runtag, mode ="FDR", main = main)
  cat("FDP at bonferroni corrected p = 0.05: ", 1 - mean(df[df$PVALUE < 0.05 /dim(df)[1], "ifcausal"]))
  return(fig)
}

scatter_plot_PIP_p <- function(pipfs, gwasfs, pipformat = "susie", main ="PIP-p"){
  df <- NULL
  for (i in 1:length(pipfs)) {
    pipres <- fread(pipfs[i], header = T)
    pipres <- data.frame(pipres[pipres$type  =="gene", ])
    pipres$runtag <- i
    if (pipformat == "susie"){
       pipres <- rename(pipres, c("susie_pip" = "pip"))
    } else if (pipformat == "SER"){
       pipres <- rename(pipres, c("SERpip" = "pip"))
    }
    gwasres <- read.table(gwasfs[i], header = T)
    gwasres <- rename(gwasres, c("MARKER_ID" = "name"))
    res <- merge(gwasres, pipres, by = "name", all = T)
    res <- res[complete.cases(res),]
    df <- rbind(df, res)
  }

  df <- rename(df, c( "PVALUE" = "TWAS.p"))
  df[,"TWAS.p"] <- -log10(df[, "TWAS.p"])

  # df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1), to=c("Non causal", "Causal"))

  df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1),to=c("darkgreen", "salmon"))

  plot(df$TWAS.p, df$pip, col = df$ifcausal, main = main, xlab = "-log10(TWAS p value)", ylab = "PIP")
  # fig <- plot_ly(data = df, x = ~ TWAS.p, y = ~ pip, color = ~ ifcausal,
  #                colors = c( "salmon", "darkgreen"), type ="scatter", text = ~ paste("Name: ", paste0(runtag,":",name),
  #                                                                  "\nChr: ", chr,  "\nPos:", pos))

}

```

```{r TWAS FDP, fig.height=4, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
gwasfs <-  paste0(simdatadir, "20201001-", tags, ".exprgwas.txt.gz")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")

f1 <- caliFDP_plot(gwasfs[1:5], susieIfs2[1:5], format= "susie", main = "TWAS FDP (low power)")
f2 <- caliFDP_plot(gwasfs[6:10], susieIfs2[6:10], format= "susie", main ="TWAS FDP (high power)")

gridExtra::grid.arrange(f1, f2, ncol =2)
```

# Parameter estimation results

Results: Each row shows parameter estimation results from 5 simulation runs with similar settings (i.e. pi1 and PVE for genes and SNPs). each row has two plots, one for gene pi1 estimation, one for enrichment (gene pi1/snp pi1). Results from each run were represented by one dot, dots with the same color come from the same run. horizontal dash lines: simulation truth,  `susietruth`, the truth in selected regions that were used to run susie iteractively (susieI).

```{r param functions}
show_param <- function(phenofs, susieIfs, susieIfs2){
  pars <- do.call(rbind, lapply(phenofs, function(x) {load(x); 
    c(phenores$param$pve.gene.truth,
      phenores$param$pve.snp.truth,
      length(phenores$batch[[1]]$param$idx.cgene)/phenores$batch[[1]]$param$J,
      length(phenores$batch[[1]]$param$idx.cSNP)/phenores$batch[[1]]$param$M)}))

  colnames(pars) <- c("PVE.gene_truth", "PVE.SNP_truth", "pi1.gene_truth", "pi1.SNP_truth")

  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec[prior.gene_rec!=0], 1), tail(prior.SNP_rec[prior.SNP_rec!=0],1))}))

  param.s.truth <- do.call(rbind, lapply(susieIfs2, function(x) {
    a <- fread(x, header = T);
    c(nrow(a[a$ifcausal == 1 & a$type == "gene" ])/ nrow(a[a$type == "gene"]),
      nrow(a[a$ifcausal == 1 & a$type == "SNP"])/ nrow(a[a$type == "SNP"]))
    }))

  pars.s <- cbind(param.s.truth, param.s)[, c(1,3,2,4)]
  colnames(pars.s) <-  paste(rep(c("pi1.gene_", "pi1.SNP_"), each = 2), c("susietruth", "susieI"), sep = "")

  df <- cbind(tags, format(pars, digits = 4), format(pars.s, digits =4))
  rownames(df) <- NULL
  return(df)

  # df %>% 
  # kable("html", escape = F) %>%
  # kable_styling("striped", full_width = F) %>%
  #  row_spec(c(1:5, 11:15), background = "#FEF3B9") %>%
  # scroll_box(width = "100%", height = "600px", fixed_thead = T)
}

plot_param <- function(df, ...){
  df <- apply(df[ , 2:ncol(df)], 2, function(x) as.numeric(x))
  st <-  cbind(df[,"pi1.gene_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "gene pi1", xaxt = "n", xlab="", xlim = c(0.8, 3.5), frame.plot=FALSE, ylim = c(0, max(dfp[,1],t) *1.05), ...)
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=2:3, 0, labels = c( "susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h=t, lty = 2, col= "salmon", lwd=1.5)
  grid()

  st <-  cbind(df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.SNP_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.SNP_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "SNP pi1", xaxt = "n", xlab="", xlim = c(0.8, 3.5), frame.plot=FALSE, ylim = c(0, max(dfp[,1],t) *1.05), ...)
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=2:3, 0, labels = c( "susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h=t, lty = 2, col= "salmon", lwd=1.5)
  grid()

  st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"]/df[,"pi1.SNP_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]/df[1,"pi1.SNP_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, min(max(dfp[,1],t) *1.05, 150)))
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=2:3, 0, labels = c("susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h= t, lty = 2, col= "darkgreen", lwd=1.5)
  grid()
}

gpip_dist <- function(susiefs, ...){
  dflist <- list()
  for (f in susiefs){
    dflist[[f]] <- read.table(f, header =T , stringsAsFactors = F)
  }
  df <- do.call(rbind, dflist)
  hist(df[df$type == "gene", "susie_pip"], xlab = "gene susie PIP", 
       breaks = 50, ylim = c(0,20), xlim=c(0,1), col = "salmon", ...)
}
```

## susieI (1)
* Regions: all regions, 500kb uniform regions.
* Susie run parameters: `L=1`.  We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). Prior variance and residual variance were calculated by SUSIE for each region.

```{r param1susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
gwasfs <-  paste0(simdatadir, "20201001-", tags, ".exprgwas.txt.gz")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

```{r pip1susie, fig.height=4, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP (low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
gridExtra::grid.arrange(f1, f2, ncol =2)
```

```{r scatter1susie, fig.height=4, fig.width=8}
par(mfrow=c(1,2))
f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "susie", main = "low power")
f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "susie", main = "high power")
```

## susieI (2)
* Regions: all regions, LD-defined regions.
* Susie run parameters: `L=1`.  We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). Prior variance and residual variance were calculated by SUSIE for each region.

```{r param6susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config6.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config6.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

## susieI (3)
* Regions: all regions, LD-defined regions.
* Susie run parameters: `L=10`.  We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). Prior variance and residual variance were calculated by SUSIE for each region.

```{r param8susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config8.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config8.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

##  Single effect model/EM (1)
* Regions: all regions, 500kb uniform regions.
* Run parameters: We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). We use the true prior variance for genes and SNPs.

```{r param1, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.SERIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.SERI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

```{r pip1ser, fig.height=4, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
gridExtra::grid.arrange(f1, f2, ncol =2)
```

```{r scatter1ser, fig.height=4, fig.width=10}
par(mfrow=c(1,2))
f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
```

##  Single effect model/EM (2)
* Regions: all regions, defined based on LDetect.
* Run parameters: We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). We use the true prior variance for genes and SNPs.

```{r param6, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config6.SERIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config6.SERI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

```{r pip6ser, fig.height=4, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
gridExtra::grid.arrange(f1, f2, ncol =2)
```

```{r scatter6ser, fig.height=4, fig.width=10}
par(mfrow=c(1,2))
f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
```

##  Single effect model/EM (3)
* Regions: regions with at most 1 causal signal, regions are defined based on LDetect.
* Run parameters: We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). We use the true prior variance for genes and SNPs.

```{r param7, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config7.SERIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config7.SERI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

```{r pip7ser, fig.height=4, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="SER", main = "PIP (low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="SER", main = "PIP(high power)" )
gridExtra::grid.arrange(f1, f2, ncol =2)
```

```{r scatter7ser, fig.height=4, fig.width=10}
par(mfrow=c(1,2))
f1 <- scatter_plot_PIP_p(susieIfs2[1:5], gwasfs[1:5], pipformat = "SER", main = "low power")
f2 <- scatter_plot_PIP_p(susieIfs2[6:10], gwasfs[6:10], pipformat = "SER", main = "high power")
```