### 2021-08-09 analysis/Paper_figures_simulation.Rmd
paper figure, examples rerun ctwas with full SNPs

---
title: "Paper figures (Simulation)"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup dir}
library(ctwas)
library(data.table)
source("~/causalTWAS/causal-TWAS/analysis/summarize_basic_plots.R")
source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
source("~/causalTWAS/causal-TWAS/analysis/ld.R")

outputdir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416/"
comparedir = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/"
runtag = "ukb-s80.45-adi"
configtag = 1
pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45_pgenfs.txt"
ld_pgenfn = "/home/simingz/causalTWAS/ukbiobank/ukb_pgen_s80.45/ukb-s80.45.2_pgenfs.txt"
exprfn = "/home/simingz/causalTWAS/simulations/simulation_ctwas_rss_20210416//ukb-s80.45-adi.expr.txt"

ld_pgenfs <- read.table(ld_pgenfn, header = F, stringsAsFactors = F)[,1]
pgenfs <- read.table(pgenfn, header = F, stringsAsFactors = F)[,1]
pvarfs <- sapply(pgenfs, prep_pvar, outputdir = outputdir)
pgens <- lapply(1:length(pgenfs), function(x) prep_pgen(pgenf = pgenfs[x],pvarf = pvarfs[x]))
exprfs <- read.table(exprfn, header = F, stringsAsFactors = F)[,1]
exprvarfs <- sapply(exprfs, prep_exprvar)

n <- pgenlibr::GetRawSampleCt(pgens[[1]])
p <- sum(unlist(lapply(pgens, pgenlibr::GetVariantCt))) # number of SNPs
J <- 8021 # number of genes

colorsall <- c("#7fc97f", "#beaed4", "#fdc086")
```

## Parameter estimation
### Parameter vs. true value

In our simulations, the SNP PVE is always set to 0.5 in different settings. The gene PVE is shown as in figures. The number of causal SNPs is always set to 2.5e * $10^{-4}$. In the two settings shown below, number of samples is 45k. For other details about our simulation settings and procedures, please see [here](simulation-ctwas-ukbWG-gtex.adipose_s80.45_041621.html). Note, setting 1 is the high power setting and setting 2 is a low power setting, I will change the setting names manually later.

Each plot show one parameter: pi.gene, pi.gene/pi.SNP (enrichment), effectsize.gene, PVE.gene. Horizontal bar shows mean true values across the 5 simulations with similar setting parameters. The results by ctwas for each simulation is shown by dots.

```{r paramplotfunc}

plot_single <- function(mtxlist, truecol, estcol, ...){
  truth <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mean(mtxlist[[x]][, truecol]))))
  est <- do.call(rbind, lapply(1:length(mtxlist), function(x) cbind(x, mtxlist[[x]][, estcol])))

  col = est[,1]
  est[,1] <- jitter(est[,1])

  plot(est, pch = 19, xaxt = "n", xlab="" ,frame.plot=FALSE, col = colorsall[col], ...)
  axis(side=1, at=1:2, labels = c("setting 1", "setting 2"), tick = F)
  #text(x=1:length(mtxlist), 0, labels = paste0("temp",1:length(mtxlist)), xpd = T, pos =1)
  for (t in 1:nrow(truth)){
    row <- truth[t,]
    segments(row[1]-0.2, row[2] , row[1] + 0.2, row[2],
       col = colorsall[t], lty = par("lty"), lwd = 2, xpd = FALSE)
  }
  grid()
}

plot_par <- function(configtag, runtag, simutaglist){
  mtxlist <- list()
  for (group in 1:length(simutaglist)){
    simutags <- simutaglist[[group]]
    source(paste0(outputdir, "config", configtag, ".R"))
    phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
    susieIfs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".s2.susieIrssres.Rd")
    susieIfs2 <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".s2.susieIrss.txt")
    mtxlist[[group]] <- show_param(phenofs, susieIfs, susieIfs2, thin = thin)
  }

  # pi1 
  par(mfrow=c(1,4))
  plot_single(mtxlist, truecol = "pi1.gene_truth", estcol = "pi1.gene_est", ylab ="gene pi1", ylim = c(0,0.02), xlim = c(0.8,2.4))
  plot_single(mtxlist, truecol = "enrich_truth", estcol = "enrich_est",ylab ="gene enrichment", ylim = c(0,120), xlim = c(0.8,2.4) )
  plot_single(mtxlist, truecol = "sigma.gene_truth", estcol = "sigma.gene_est", ylab = "gene effect size", ylim = c(0.01, 0.03), xlim = c(0.8,2.4))
  plot_single(mtxlist, truecol = "PVE.gene_truth", estcol = "PVE.gene_est", ylab ="gene PVE", ylim = c(0, 0.1), xlim = c(0.8,2.4))

  return(mtxlist)
}

```

Figure (supplementary):
```{r paramplot, fig.width= 8, fig.height= 3}
simutaglist = list(paste(4, 1:5, sep = "-"), paste(10, 1:5, sep="-"))
mtxlist <- plot_par(configtag, runtag, simutaglist)
```

Back up figures

I have in total 10 x 2 = 20 settings to use generate the parameter figures.

### Comparison of PVE.gene vs. MESC estimates.
Leave it to Kevin.

## PIP calibration
### PIP plot under different settings.

```{r pipplotfunc}
plot_PIP <- function(configtag, runtag,  simutags, ...){
   phenofs <- paste0(outputdir, "ukb-s80.45-adi", "_simu", simutags, "-pheno.Rd")
   susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")

   f1 <- caliPIP_plot(phenofs, susieIfs, ...) 
   return(f1)
}
```

```{r pipplot, fig.width= 8, fig.height= 4}
simutaglist = list(paste(4, 1:5, sep = "-"), paste(10, 1:5, sep="-"))
f1 <- plot_PIP(configtag, runtag, simutaglist[[1]], main = "high gene PVE")
f2 <-  plot_PIP(configtag, runtag, simutaglist[[2]], main = "low gene PVE")
gridExtra::grid.arrange(f1, f2, ncol =2)
```

### Comparison with other methods
Bar plot: each bar shows the number of genes, colored by causal status. Use a different color for each method. The method and cut off values:
* ctwas: PIP 0.8
* FUSION fdr: 0.05
* FUSION bonferroni: 0.05
* COLOC PP4: 0.8
* FOCUS PIP: 0.8
* SMR FDR: 0.05
* SMR HEIDI: HEIDI p > 0.05, SMR FDR < 0.05

Multiple bar plots, different settings: high gene power and low gene power.
```{r ncausalplotfunc}

get_ncausal_df <- function(pfiles, cau, cut = 0.8, 
                           method = c("ctwas", "fusionfdr", "fusionbon","coloc", "focus", "smr", "smrheidi")){
  df <- NULL
  for (i in 1:length(pfiles)) {
    res <- fread(pfiles[i], header = T)
    # res <- res[complete.cases(res),]
    if (method == "ctwas"){
        res <- data.frame(res[res$type  =="gene", ])
        res$ifcausal <- ifelse(res$id %in% cau[[i]], 1, 0)
        res <- res[res$susie_pip > cut,]
    } else if (method == "fusionfdr"){
        res$FDR <- p.adjust(res$TWAS.P, method = "fdr")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
    } else if  (method == "fusionbon"){
        res$FDR <- p.adjust(res$TWAS.P, method = "bonferroni")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
    } else if  (method == "coloc"){
        res <- res[res$COLOC.PP4 > cut,]
        res$ifcausal <- ifelse(res$ID %in% cau[[i]], 1, 0)
    } else if (method == "focus"){
        res <- res[res$mol_name != "NULL",]
        res <- res[res$pip > cut, ] 
        res$ifcausal <- ifelse(res$mol_name %in% cau[[i]], 1, 0)
    } else if (method == "smr"){
        res$FDR <- p.adjust(res$p_SMR, method = "fdr")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
    } else if (method == "smrheidi"){
        res <- res[res$p_HEIDI > 0.05,]
        res$FDR <- p.adjust(res$p_SMR, method = "fdr")
        res <- res[res$FDR < cut,]
        res$ifcausal <- ifelse(res$probeID %in% cau[[i]], 1, 0)
    } else{
      stop("no such method")
    }

    df.rt <- rbind(c(nrow(res[res$ifcausal == 0, ]), 0, i),
                   c(nrow(res[res$ifcausal == 1, ]), 1, i))
    df <- rbind(df, df.rt)
  }
  colnames(df) <- c("count", "ifcausal", "runtag")
  df <- data.frame(df)
  df$method <- method
  return(df)
}

plot_ncausal <- function(configtag, runtag,  simutags, colors, ...){
  phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
  cau <- lapply(phenofs, function(x) {load(x);get_causal_id(phenores)})

  susieIfs <- paste0(outputdir, runtag, "_simu",simutags, "_config", configtag,".susieIrss.txt")
  fusioncolocfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.coloc.result")
  focusfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.focus.tsv")
  smrfs <- paste0(comparedir, runtag, "_simu", simutags, ".Adipose_Subcutaneous.smr")

  ctwas_df <- get_ncausal_df(susieIfs, cau= cau, cut = 0.8, method ="ctwas")
  fusionfdr_df <- get_ncausal_df(fusioncolocfs, cau= cau, cut = 0.05, method = "fusionfdr")
  fusionbon_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.05, method = "fusionbon")
  coloc_df <- get_ncausal_df(fusioncolocfs , cau= cau, cut = 0.8, method = "coloc")
  focus_df <- get_ncausal_df(focusfs , cau= cau, cut = 0.8, method = "focus")
  smr_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smr")
  smrheidi_df <- get_ncausal_df(smrfs, cau= cau, cut = 0.05, method = "smrheidi")

  df <- rbind(ctwas_df, fusionfdr_df, fusionbon_df, coloc_df, focus_df, smr_df, smrheidi_df)
  df$ifcausal <- df$ifcausal + as.numeric(as.factor(df$method))*10
  df$ifcausal <- as.factor(df$ifcausal)
  fig <- ggbarplot(df, x = "method", y = "count", add = "mean_se", fill = "ifcausal", palette = colors, legend = "none", ...) + grids(linetype = "dashed")
  fig
}

```

```{r ncausalplot, fig.width= 10, fig.height= 4}
colset = c("#ebebeb", "#bebada" , "#ebebeb", "#fb8072", "#ebebeb","#ffffb3","#ebebeb", "#8dd3c7", "#ebebeb","#8dd3c7", "#ebebeb", "#87CEFA", "#ebebeb","#87CEFA") # coloc, ctwas, focus, fusionbon, fusionfdr, smrheidi, smr
f1 <- plot_ncausal(configtag, runtag,  simutaglist[[1]], colors = colset, ylim= c(0,150), main = "high gene PVE")
f2 <- plot_ncausal(configtag, runtag,  simutaglist[[2]], colors = colset, ylim= c(0,150), main = "low gene PVE")
gridExtra::grid.arrange(f1, f2, ncol=2)
```

## Examples
### FP example 1
cTWAS avoids the FP error. In this case, the false positive gene (from TWAS) is caused by LD of eQTLs with a causal gene's eQTLs.

```{r regionplotfunc}
plot_region <- function(runtag, simutag, configtag, chr, startpos = NULL, endpos = NULL, rerun_ctwas = F, plot_ceqtl = F,
                        plot_eqtl = F){
  pf <- paste0(outputdir, runtag, "_simu",simutag)
  phenof <- paste0(outputdir, runtag, "_simu", simutag, "-pheno.Rd")

  b1 <- fread(paste0(pf, ".snpgwas.txt.gz"), header =T)
  setnames(b1, old = "pos", new = "p0")
  b2 <- fread(paste0(pf, ".exprgwas.txt.gz"), header =T)
  b <- rbind(b1, b2, fill = T)
  b <- b[b$chrom == chr,]

  susief <- paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".susieIrss.txt")
  a <- fread(susief, header =T)
  a <- a[a$chrom == chr,]

  if (!is.null(startpos)){
    a <- a[a$pos > startpos & a$pos < endpos]
    b <- b[b$p0 > startpos & b$p0 < endpos]
  }

  a <- merge(a, b, by = "id", all = T)
  a[is.na(a$type),"type"] <- "SNP"

  if (isTRUE(rerun_ctwas)){
      ld_exprfs <- paste0(outputdir, runtag, "_simu", simutag, "_chr", 1:22, ".expr.gz")
      load(paste0(outputdir, runtag, "_simu", simutag, "_config", configtag, ".s2.susieIrssres.Rd"))
      source(paste0(outputdir, "config", configtag, ".R"))
      group_prior <- group_prior_rec[, ncol(group_prior_rec)] 
      group_prior[2] <- group_prior[2] * thin
      group_prior_var <- group_prior_var_rec[, ncol(group_prior_var_rec)]
      temp_reg <- data.frame("chr" = paste0("chr",chr), "start" = startpos, "stop" = endpos)
      write.table(temp_reg, file= "temp_reg.txt" , row.names=F, col.names=T, sep="\t", quote = F)
      z_gene <-  a[a$type == "gene", c("id", "t.value")]
      colnames(z_gene) <- c("id", "z")
      z_snp <-  a[a$type == "SNP", c("id", "alt", "ref", "t.value")]
      colnames(z_snp) <- c("id", "A1", "A2", "z")
      ctwas_rss(z_gene, z_snp, ld_exprfs= ld_exprfs, ld_pgenfs = ld_pgenfs, ld_R_dir = NULL, ld_regions_custom = "temp_reg.txt", thin = 1, outputdir = ".", outname = "temp", ncore = 1, ncore.rerun = 1, prob_single = 0,  group_prior = group_prior, group_prior_var = group_prior_var, estimate_group_prior = F, estimate_group_prior_var = F, recover_strand_ambig_z = F)
      a2 <- fread("temp.susieIrss.txt", header = T)
      a <- merge(a2, b, by = "id", all = T)
      # file.remove("temp.susieIrss.txt")
      # file.remove("temp.temp.susieIrssres.Rd")
      # file.remove("temp.regions.txt")
      # file.remove("temp_reg.txt")
   }

  load(phenof)
  cau <- get_causal_id(phenores)

  a$ifcausal <- ifelse(a$id %in% cau, 1, 0)
  a[is.na(a$type),"type"] <- "SNP"

  a[, "PVALUE"] <- -log10(a[, "PVALUE"])
  a$r2max <- get_ld2(ids =a$id, phenores = phenores, pgenfs = pgenfs, exprfs = exprfs, chrom = chr)

  r2cut <- 0.4

  layout(matrix(1:2, ncol = 1), widths = 1, heights = c(1.5,1.5), respect = FALSE)
  par(mar = c(0, 4.1, 4.1, 2.1))
  plot(a[a$type=="SNP"]$p0, a[a$type == "SNP"]$PVALUE, pch = 19, xlab="Genomic position" ,frame.plot=FALSE, col = "white", ylim= c(-0.1,1.1), ylab = "ctwas PIP", xaxt = 'n')
  grid()
  points(a[a$type=="SNP"]$p0, a[a$type == "SNP"]$susie_pip, pch = 21, xlab="Genomic position", bg = colorsall[1])
  points(a[a$type=="SNP" & a$r2max > r2cut, ]$p0, a[a$type == "SNP"  & a$r2max >r2cut]$susie_pip, pch = 21, bg = "purple")
  points(a[a$type=="SNP" & a$ifcausal == 1, ]$p0, a[a$type == "SNP" & a$ifcausal == 1]$susie_pip, pch = 21, bg = "salmon")
  points(a[a$type=="gene" ]$p0, a[a$type == "gene" ]$susie_pip, pch = 22, bg = colorsall[1], cex = 2)
  points(a[a$type=="gene" & a$r2max > r2cut, ]$p0, a[a$type == "gene"  & a$r2max > r2cut]$susie_pip, pch = 22, bg = "purple", cex = 2)
  points(a[a$type=="gene" & a$ifcausal == 1, ]$p0, a[a$type == "gene" & a$ifcausal == 1]$susie_pip, pch = 22, bg = "salmon", cex = 2)
  if (isTRUE(plot_ceqtl)){
      for (cgene in a[a$type=="gene" & a$ifcausal == 1, ]$id){
     load(paste0(tools::file_path_sans_ext(exprfs[a[a$id == cgene, "chrom.x"][[1]]]), "qc.Rd"))
     eqtls <- rownames(wgtlist[[cgene]])
     points(a[a$id %in% eqtls,]$p0, rep( -0.15, nrow(a[a$id %in% eqtls,])), pch = "|", col = "salmon", cex = 1.5)
    }
  }

  if (isTRUE(plot_eqtl)){
    for (cgene in a[a$type=="gene" & a$PVALUE >5, ]$id){
     load(paste0(tools::file_path_sans_ext(exprfs[a[a$id == cgene, "chrom.x"][[1]]]), "qc.Rd"))
     eqtls <- rownames(wgtlist[[cgene]])
     points(a[a$id %in% eqtls,]$p0, rep( -0.15, nrow(a[a$id %in% eqtls,])), pch = "|", col = "black", cex = 1.5)
    }
  }

  legend(min(a$p0), y= 1.3 ,c("Gene", "SNP"), pch = c(22,21), title="shape legend", bty ='n', cex =0.8, title.adj = 0)
  legend(min(a$p0), y= 0.8 ,c("Causal", "Noncausal, R2 > 0.4", "Noncausal, R2 <= 0.4"), pch = 19, col = c("salmon", "purple", colorsall[1]), title="color legend", bty ='n', cex =0.8, title.adj = 0)
  par(mar = c(4.1, 4.1, 0.5, 2.1))
  plot(a[a$type=="SNP"]$p0, a[a$type == "SNP"]$PVALUE, pch = 21, xlab="Genomic position" ,frame.plot=FALSE, bg = colorsall[1], ylab = "TWAS -log10(p value)", panel.first = grid(), ylim =c(0, max(a$PVALUE)*1.2))
  points(a[a$type=="SNP" & a$r2max > r2cut ]$p0, a[a$type == "SNP"  & a$r2max > r2cut]$PVALUE, pch = 21, bg = "purple")
  points(a[a$type=="SNP" & a$ifcausal == 1, ]$p0, a[a$type == "SNP" & a$ifcausal == 1]$PVALUE, pch = 21, bg = "salmon")
  points(a[a$type=="gene" ]$p0, a[a$type == "gene" ]$PVALUE, pch = 22, bg = colorsall[1], cex = 2)
  points(a[a$type=="gene" & a$r2max > r2cut, ]$p0, a[a$type == "gene"  & a$r2max > r2cut]$PVALUE, pch = 22, bg = "purple", cex = 2)
  points(a[a$type=="gene" & a$ifcausal == 1, ]$p0, a[a$type == "gene" & a$ifcausal == 1]$PVALUE, pch = 22, bg = "salmon", cex = 2)
  abline(h=-log10(0.05/J), col ="red", lty = 2)

  return(a)
}
```

```{r regionplot1, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "4-4"
a <- plot_region(runtag, simutag, configtag, chr = 4, startpos = 43965045, endpos = 45189157,  plot_ceqtl = T, rerun_ctwas = T)
```

### FP example 2
cTWAS avoids the FP error. In this case, the false positive gene (from TWAS) is caused by LD of eQTLs with a causal SNP nearby.

```{r regionplot2, echo=FALSE,results='hide',fig.keep='all', message=F}
runtag = "ukb-s80.45-adi"
simutag <- "3-2"         
a <- plot_region(runtag, simutag, configtag, chr = 19, startpos = 16824416, endpos = 17559862, rerun_ctwas = T, plot_eqtl = T)
```

### TP example
cTWAS is able to find true positives. This gene has one eQTL, cTWAS choose this gene because it uses a  prior favoring genes, it didn't reach significance level by TWAS after bonferron correction.

```{r regionplot3, echo=FALSE,results='hide',fig.keep='all', message=F}
simutag <- "4-4"
a <- plot_region(runtag, simutag, configtag, chr = 1, startpos = 37549183 , endpos =38731847, rerun_ctwas = T)
```