### 2023-03-22 analysis/simulation-ctwas-ukbWG-gtex.adipose_visual.Rmd

---
title: "Visualize ukb whole genome, using ctwas results"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message= F, warning = F)
```

```{r functions file}
source("~/causalTWAS/causal-TWAS/analysis/summarize_ctwas_plots.R")
source("~/causalTWAS/causal-TWAS/analysis/visual_twas_plots.R")
library(plyr)

plot_all <- function(susief, pf, phenof, colocf = NULL, chr, ldres = NULL, thin = 1){
  a <- fread(susief, header =T)
  setnames(a, old = "id", new = "name")
  b1 <- fread(paste0(pf, ".snpgwas.txt.gz"), header =T)
  b2 <- fread(paste0(pf, ".exprgwas.txt.gz"), header =T)
  setnames(b1, old = "pos", new = "p0")
  b <- rbind(b1, b2, fill = T)
  setnames(b, old = c('PVALUE', 'id'),
         new = c('pvalue', 'name'))

  load(phenof)
  cau <- get_causal_id(phenores)

  a <- a[a$chrom == chr,]
  b <- b[b$chrom == chr,]

  a <- merge(a, b, by = "name", all = T)
  a$ifcausal <- ifelse(a$name %in% cau, "Causal", "Non causal" )
  a[is.na(a$type),"type"] <- "SNP"

  if (thin !=1){
    a <- a[ !is.na(a$type) | a$ifcausal == "Causal" | 
             (a$ifcausal == "Non causal" & rbinom(nrow(a), 1, thin) ==1)]
  }

  if (!is.null(ldres)){
    ld <- read.table(ldres, header =T)
    a <- merge(a, ld, by = "name", all.x = T)}

  colnames(a)[colnames(a) == "susie_pip"] <-  "susiePIP"

  a[, "PVALUE"] <- -log10(a[, "pvalue"])

  dt.gene <- a[a$type == "gene", ]
  dt.snp <- a[a$type == "SNP", ]

  if (!is.null(colocf)) {
    a <- fread(colocf, header = T)
    setnames(a, old = "ID", new = "name")
    dt.gene <- merge(dt.gene, a, by ="name")
  }

  fig1 <- plot_susie_pip(dt.gene)  # %>%  add_trace(p, x = list(1), y = list(1), name = " ", showlegend = F)
  # fig2 <- plot_lbf(dt.gene)
  fig3 <- plot_p(dt.gene)

  fig4 <- plot_susie_pip(dt.snp)
  # fig5 <- plot_lbf(dt.snp) 
  fig6 <- plot_p(dt.snp) %>% layout( plot_bgcolor='#EFF6FC')

  if (is.null(colocf)) {
     fig <- subplot(fig1, fig3, fig4, fig6, nrows=4, shareX = T, titleY = T)
  } else {
     fig7 <- plot_coloc_pp4(dt.gene)
     fig <- subplot(fig1, fig3, fig4, fig6, fig7, nrows=5, shareX = T, titleY = T)

  }

  fig <- fig %>% layout(autosize = F, width = 600, height = 500, margin = 0.02,
                        xaxis = list(title = "Genomic position"))

  fig <- fig %>% toWebGL()
  fig
}
```

## Scatter plot

```{r sim 2, out.width = "98%", out.height="95%"}
outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210416/"
runtag <- "ukb-s80.45-adi"
simutags <- paste(3, 1:5, sep = "-")
configtag <- 1

susiefs <- paste0(outputdir, runtag, "_simu", simutags, "_config", configtag, ".susieIrss.txt")
pfs <- paste0(outputdir, runtag, "_simu",simutags)
gwasfs <- paste0(pfs, ".exprgwas.txt.gz")
phenofs <- paste0(outputdir, runtag, "_simu", simutags, "-pheno.Rd")
scatter_plot_PIP_p(phenofs, susiefs, gwasfs, main ="PIP-p")
```

## Zoom view
FP example, LD with causal gene, GNPDA2
under 4-1 to 4-5
```{r sim 2, out.width = "98%", out.height="95%"}
chrom <- 4
runtag <- 4
plot_all(susiefs[runtag], pfs[runtag], phenofs[runtag], chrom, thin = 0.01)
```

FP example: LD with causal SNP, GALNT10
```{r sim 2, out.width = "98%", out.height="95%"}
runtag <- 5
chrom <- 1
plot_all(susiefs[runtag], pfs[runtag], phenofs[runtag], chrom, thin = 0.01)
```

under 10-1 to 10-5
FP example: LD with causal SNP
```{r sim 2, out.width = "98%", out.height="95%"}
runtag <- 5
chrom <- 1
plot_all(susiefs[runtag], pfs[runtag], phenofs[runtag], chrom, thin = 0.01)
```

True positive example: c1orf109
```{r sim 2, out.width = "98%", out.height="95%"}
runtag <- 4
chrom <- 1
plot_all(susiefs[runtag], pfs[runtag], phenofs[runtag], chrom, thin = 0.01)
```

# region plot with coloc results
```{r sim 2, out.width = "98%", out.height="95%"}
runtag <- "ukb-s80.45-adi"
i <- 4
chrom <- 5
comparedir <- "/project2/mstephens/causalTWAS/simulations/simulation_ctwas_rss_20210416_compare/"
colocfs <- paste0(comparedir, runtag, "_simu", simutags,".Adipose_Subcutaneous.coloc.result")
plot_all(susiefs[i], pfs[i], phenofs[i], chrom, colocf = colocfs[i], thin = 0.01)
```